---
title: "Adding Categories for Conceptual Model"
author: "Erika W"
date: "3/9/2021"
output:
  html_document:
    code_download: true
    keep_md: true
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=F, message=F}
rm(list = ls())
library(data.table)
library(lubridate)
library(sf)
library(tidyverse)
```

## Intro

This rmd produces a limited combined dataset, retaining only those analytes identified within our conceptual model, and including a new column "SelectionList" by which analytes can be subsetted according to their role in the conceptual model.

There are two approaches that should work - whichever is most compatible with needs for later updates, or future documentation. 

## Data Splitting

### Method 1: Lists

If it's most simple to have a transparent list of analytes included in each category, documented within the R code, this approach will work. The code is a little longer. 

#### Load Data

Start with merged CEDEN and SURF data, with duplicates found and removed per the QA/QC protocol outlined at: https://github.com/WWU-IETC-R-Collab/CEDENSURF-mod/blob/main/CEDENSURF_Merge.md

```{r}
# Load CEDEN Data
CEDENSURF <- fread("https://github.com/WWU-IETC-R-Collab/CEDENSURF-mod/raw/main/Data/Output/CEDENSURFMod.csv")

# Create empty column to fill with values

CEDENSURF$SelectList <- NA

# Keep a record (DF) of unique analytes?

AnalyteList <- sort(unique(CEDENSURF$Analyte)) #1053 records
```

#### Append from Selection Lists

```{r}
## Used grepl() to help find all Analytes that matched the conceptual model, ie: AnalyteList[(grepl("nitr", AnalyteList))

## Water Quality Parameters
            
WQP_SelectList <- c("temperature", "ph", "oxygen",
                    "electricalconductivity", 
                    "nitrate as n", "nitrite as n", "nitrogen",
                    "nitrate + nitrite as n", "phosphorus as p",
                    "turbidity","settleable solids", 
                    "suspended sediment concentration", 
                    "salinity","sodium")


CEDENSURF$SelectList[CEDENSURF$Analyte %in% WQP_SelectList] <- "WQP"

# Metals

Metal_SelectList <- c("mercury", "selenium") # # Only one record of mercury (ii)r; chose to omit. # did not find methylmercury in list. 
# Others may want? "silver", 


CEDENSURF$SelectList[CEDENSURF$Analyte %in% Metal_SelectList] <- "Metal"

## Organophosphates

ORG_SelectList <- c("malathion", 
                    "diazinon", "diazinon degradate", 
                    "diazinon oxon", "diazoxon",
                    "chlorpyrifos")

CEDENSURF$SelectList[CEDENSURF$Analyte %in% ORG_SelectList] <- "OrganoP"

## Neonicotinoids

Neon_SelectList <- c("hydroxy-imidacloprid", "imidacloprid")

CEDENSURF$SelectList[CEDENSURF$Analyte %in% Neon_SelectList] <- "Neon"

## Pyrethroids

Pyre_SelectList <- c("binfenthrin", "cyfluthrin", 
                     "esfenvalerate", "esfenvalerate/fenvalerate",
                     "esfenvalerate/fenvalerate-1",
                     "esfenvalerate/fenvalerate-2") # did not find any binfenthrin

CEDENSURF$SelectList[CEDENSURF$Analyte %in% Pyre_SelectList] <- "Pyrethroids"

## GABA inhibitors

GABA_SelectList <- AnalyteList[grepl('fipronil', AnalyteList)] # THIS IS THE SAME AS: <- c("desulfinyl fipronil", "desulfinyl fipronil amide", "fipronil", "fipronil amide", "fipronil desulfinyl", "fipronil desulfinyl amide", "fipronil detrifluoromethylsulfinyl", "fipronil sulfide","fipronil sulfone")

CEDENSURF$SelectList[CEDENSURF$Analyte %in% GABA_SelectList] <- "GABA"

## Glyphosate (it's only one!)

CEDENSURF$SelectList[CEDENSURF$Analyte == "glyphosate"] <- "Glyphosate"

## Atrazine

Atraz_SelectList <-  c("atrazine", "atrazine degradate")
# Chose to omit single records: "desethyl-atrazine""desisopropyl-atrazine", "hydroxyatrazine","hydroxyatrazine, 2-" )

CEDENSURF$SelectList[CEDENSURF$Analyte %in% Atraz_SelectList] <- "Atrazine"

```

#### Save categories assigned to analytes

```{r}
## Save Analyte Table showing assigned categories

AnalyteTable <- CEDENSURF %>% select(Analyte, SelectList) %>% distinct(Analyte, .keep_all = T) # 44 records

write.csv(x = AnalyteTable, 
          file = "Data/Output/CEDENSURF_AnalytesCategorized.csv", 
          na = "", row.names = F)
```

### Save Categorized, Limited CEDENSURF 

```{r}
## Remove rows with irrelevant Analytes?
CEDENSURF<- CEDENSURF %>% filter(!is.na(SelectList))
```

The result is `r nrow(CEDENSURF)` records, all appended with appropriate selection categories according to the conceptual model

```{r}
head(CEDENSURF %>% select(Date, Analyte, Result, Unit, StationName, SelectList))
```

```{r}
write.csv(x = CEDENSURF, file = "Data/Output/CEDENSURF_Limited.csv", 
          row.names = F)
```


### Method 2: Table

#### Load Data

Two files:

1. Merged CEDEN and SURF data, with duplicates found and removed per the QA/QC protocol outlined at: https://github.com/WWU-IETC-R-Collab/CEDENSURF-mod/blob/main/CEDENSURF_Merge.md

2. Table containing analyte names and selection categories. This would be made externally, but for this example I'll use a "mock table" adapted from my output above. 

A suggested structure of this table is:
```{r}
# Load external table (Mock Data)

AnalyteTable <- CEDENSURF %>% select(Analyte, SelectList) %>% distinct(Analyte, .keep_all = T) # 44 records

head(AnalyteTable)

# Load CEDEN Data
CEDENSURF2 <- fread("https://github.com/WWU-IETC-R-Collab/CEDENSURF-mod/raw/main/Data/Output/CEDENSURFMod.csv")
```

With this, we could append the category to the original data using a merge

```{r}
CEDENSURF2 <- merge(x = CEDENSURF2, y = AnalyteTable) # removes rows of x with no match in Y

head(CEDENSURF2 %>% select(Date, Analyte, Result, StationName, SelectList))
```

The result is `r nrow(CEDENSURF)` records, all appended with appropriate selection categories according to the conceptual model

<br>


