---
title: "Convert to Wide"
author: "Erika W"
date: "4/5/2021"
output:
  html_document:
    code_download: true
    keep_md: true
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, warning=F, message=F}
rm(list = ls())
library(data.table)
library(lubridate)
library(sf)
library(tidyverse)
```


## Introduction

Because NETICA requires the nodes to be columns, this data needs to be transformed to wide format.

1. Start with the integrated, modified CEDEN and SURF datasets that have had the conceptual model categories appended
2. Convert units to be consistent within each analyte category (node)
3. Summarize into wide format

## 1. Load Data
```{r}
# Load data
CEDENSURF <- fread("https://github.com/WWU-IETC-R-Collab/CEDENSURF-mod/raw/main/Data/Output/CEDENSURF_Limited.csv") %>% select(Analyte, Result, Unit,CollectionMethod, Matrix, Date, Subregion, StationName, Latitude, Longitude,SelectList)
```


## 2. Convert Units & Names {.tabset}

Units should be consistent within each analyte-Matrix combination. Initially, they were not. I separated the dataframe into categories of the conceptual model, and worked within those WideSubsets to unify units of each analyte-matrix combination.

Because Netica cannot allow spaces, -, /, or other symbols in the column names, analyte names must also be converted to usable formats.

```{r}
CEDENSURF %>% group_by(Analyte, Matrix) %>%
  summarize(n_unit = n_distinct(Unit)) %>% head(.)
```

### WQP

Few discrepancies in units between analytes in each category:

1. Phosphorous; 33/900 records had mg/Kg dw rather than mg/L

2. Turbidity; One analyte is specifically turbidity, units are NTU. Some variables that may also indicate turbidity had different units, but are possibly not comparable.

No conversions were made for WQP Parameters, though some analytes were omitted.

```{r}
WQP <- CEDENSURF %>% filter(SelectList == "WQP") 
```

#### Univariate - names and units as expected

Temp - *C
pH - no units

#### Oxygen - mg/L

**water** (3504 records) mg/L
**sediment** (281 records) mg/L

89 water records were converted from %DO, assuming an average water temperature of 20C.
    *Oxygen mg/L = (% Saturation)/10.995*
    https://www.waterontheweb.org/under/waterquality/dosatcalc.html

```{r, results = 'hide'}
# Oxygen - 
# Water Units = 3,416 records in mg/L and 89 in %
# Sediment Unit = 281 records mg/L

# Convert % to mg/L

WQP$Result[WQP$Analyte == "oxygen" & WQP$Unit == "%"] <- WQP$Result[WQP$Analyte == "oxygen" & WQP$Unit == "%"] /10.995

# Correct units

WQP$Unit[WQP$Analyte == "oxygen"] <- "mg/L"

# Check

WQP %>% filter(Analyte == "oxygen") %>% filter(Matrix == "water") %>% distinct(Unit)

WQP %>% filter(Analyte == "oxygen") %>% filter(Matrix == "water") %>% nrow(.)
```

#### Nitrogen

*Water* - mg/L
nitrate as n (379 records)
nitrite as n (444 records)

Total N:
nitrogen (709 records)
nitrate + nitrite as n (844 records)

*Sediment* - % dw
nitrogen (33 records)

```{r, results = 'hide'}
# Nitrate - Unit = mg/L (nitrate as n)

WQP %>% filter(Analyte== "nitrate as n") %>% distinct(Unit)
WQP$Analyte[WQP$Analyte == "nitrate as n"] <- "nitrate" #rename to omit spaces

# Nitrite  - Unit = mg/L (nitrite as n) 

WQP %>% filter(Analyte== "nitrite as n") %>% distinct(Unit)
WQP$Analyte[WQP$Analyte == "nitrite as n"] <- "nitrite"

# Combined as Nitrogen - 748 records; mg/L and %dw

    # Nitrogen/sediment is in % dw

    WQP %>% filter(grepl("nitro", Analyte)) %>% 
            filter(Matrix == 'sediment') %>% distinct(Unit)

    # Nitrogen/water is in mg/L
    
    WQP %>% filter(grepl("nitro", Analyte)) %>% 
            filter(Matrix == 'water') %>% distinct(Unit)
    
# Combined as Nitra + Nitrite - 844 records; mg/L (all water)

    WQP %>% filter(Analyte == "nitrate + nitrite as n") %>% 
            filter(Matrix == 'water') %>% distinct(Unit)
    WQP$Analyte[WQP$Analyte == "nitrate + nitrite as n"] <- "nitrogen" # rename

```

#### Phosphorous 

**water** (819 records) mg/L
**sediment** (27 records) mg/Kg dw

```{r, results = 'hide'}
# Phosphorous ("phosphorous as p")

# Water - 819 records (mg/L)

WQP %>% filter(grepl("phos", Analyte)) %>% 
            filter(Matrix == 'water') %>% distinct(Unit)

# Sediment - 27 records (mg/Kg dw)

WQP %>% filter(grepl("phos", Analyte)) %>% 
            filter(Matrix == 'sediment') %>% distinct(Unit)

# All named same?
WQP %>% filter(grepl("phos", Analyte)) %>% distinct(Analyte)

WQP$Analyte[WQP$Analyte == "phosphorus as p"] <- "phosphorus"
```

#### Turbidity - NTU

Units differ between three analytes that might refer to turbidity:
    # 1,600 turbidity (NTU), 11 settleable solids (mL/L/hr), 
    #94 suspended sediment concentration (mg/L)
    
I chose to remove all but Turbidity from our dataset.

```{r, results = 'hide'}
# Turbidity

    ## Units differ between three analytes that might refer to turbidity:
    # 1,600 turbidity (NTU), 11 settleable solids (mL/L/hr), 
    #94 suspended sediment concentration (mg/L)
    
    WQP %>% filter(grepl("turb", Analyte)) %>% distinct(Unit)
    
    ## Chose to remove the other two from our DF
    WQP<- WQP %>% filter(!grepl("settle", Analyte)) %>%
                  filter(!grepl("suspend", Analyte))
```

#### Salinity - ppt

Converted all to ppt

```{r, results = 'hide'}
# Salinity

WQP %>% filter(grepl("salinity", Analyte)) %>% distinct(Unit)

    # ppt and psu are equivalent measures. Convert all to ppt
    
    # Run conversion on subset of data
    Sal <- WQP %>% filter(Analyte == "salinity") %>% mutate(Unit = "ppt") 
    
    # Remove that subset from Main df
    WQP<- WQP %>% filter(!Analyte == "salinity")
    
    # Bind converted back to Main df
    WQP <- rbind(WQP, Sal)
```

#### **WQP Result**

```{r}
WQP %>%
  group_by(Analyte, Matrix, Unit) %>%
  summarise(n = n(),
            mean = mean(Result))
```


```{r}
Wide.WQP.Waterdf <- WQP %>% filter(!Matrix == "sediment") %>%
  group_by(Date, Latitude, Longitude, Analyte, Matrix) %>%
  summarize(Subregion = first(Subregion),
            Mean = mean(Result, na.rm = T)) %>%
  pivot_wider(names_from = Analyte,
              names_repair = "check_unique",
              values_from = Mean) # Values to fill columns
```


```{r}
write.csv(x = Wide.WQP.Waterdf , file = "Data/Output/WideSubsets/WQP.Wide.water.csv", 
          row.names = F)
```

<br>

### Metals

Metals were measured in ug/L and mg/Kg dw. Mercury also had few measures in ng/L and ug/Kg - these values were divided by 1000 to convert to ug/L and mg/Kg. 

**water**
mercury ug/L 
selenium (208 records) mg/L


**sediment**


mercury (mg/Kg dw
selenium (38 records) mg/Kg dw

```{r, results = 'hide'}

Metal <- CEDENSURF %>% filter(SelectList == "Metal")

# Selenium

    # Water - 208 records, mg/L
    Metal %>% filter(grepl("selenium", Analyte))  %>% 
              filter(Matrix == 'water') %>% distinct(Unit)

    # Sediment - 38 records, mg/Kg dw
    Metal %>% filter(grepl("selenium", Analyte))  %>% 
               filter(Matrix == 'sediment') %>% distinct(Unit)

# Mercury

    # Water - ug/L & ng/L
    
    Metal %>% filter(grepl("mercury", Analyte))%>% 
              filter(Matrix == 'water') %>% distinct(Unit)
    
        # Convert ng to ug
          Metal$Result[
            Metal$Analyte == "mercury" & Metal$Unit == "ng/L"] <- Metal$Result[
              Metal$Analyte == "mercury" & Metal$Unit == "ng/L"]/1000
          
        # Correct units
          Metal$Unit[Metal$Analyte == "mercury" & Metal$Matrix == "water"] <- "ug/L"
    
          
    # Sediment - (19) ug/Kg dw	& (57) mg/Kg dw
    
    Metal %>% filter(grepl("mercury", Analyte))%>% 
              filter(Matrix == 'sediment') %>% distinct(Unit)

            # Convert ug to mg
              Metal$Result[
                Metal$Analyte == "mercury" & Metal$Unit == "ug/Kg dw"] <- Metal$Result[
                  Metal$Analyte == "mercury" & Metal$Unit == "ug/Kg dw"]/1000
              
            # Correct units
              Metal$Unit[Metal$Analyte == "mercury" &
                           Metal$Matrix == "sediment"] <- "mg/Kg dw"
```


#### **Metal Result**

```{r}
Metal %>%
  group_by(Analyte, Matrix, Unit) %>%
  summarise(n = n(),
            mean = mean(Result))
```


```{r}
Wide.Metal.Waterdf <- Metal %>% filter(!Matrix == "sediment") %>%
  group_by(Date, Latitude, Longitude, Analyte, Matrix) %>%
  summarize(Subregion = first(Subregion),
            Mean = mean(Result, na.rm = T)) %>%
  pivot_wider(names_from = Analyte,
              names_repair = "check_unique",
              values_from = Mean) # Values to fill columns


write.csv(x = Wide.Metal.Waterdf , file = "Data/Output/WideSubsets/Metal.Wide.water.csv", 
          row.names = F)
```

<br>

### Organophosphates

```{r}
OrganoP <- CEDENSURF %>% filter(SelectList == "OrganoP")
```

#### chlorpyrifos - ppb

```{r, results = 'hide'}
# chlorpyrifos

  # Sediment:	ppb	331	, ng/g dw	71	(equivalent units)
    OrganoP$Unit[OrganoP$Analyte == "chlorpyrifos" &
                           OrganoP$Matrix == "sediment"] <- "ppb"

  # Water:	ng/L	220, pg/L	5, ppb	1275	
      # ppb = pg / (1000*1000)
      # ppb = ng/L / 1000
    
      # Convert ng/L to ppb
        OrganoP$Result[OrganoP$Analyte == "chlorpyrifos" &
                  OrganoP$Unit == "ng/L"] <- OrganoP$Result[
                    OrganoP$Analyte == "chlorpyrifos" &
                    OrganoP$Unit == "ng/L"] /1000
      # Convert pg/L to ppb
            OrganoP$Result[OrganoP$Analyte == "chlorpyrifos" &
                  OrganoP$Unit == "pg/L"] <- OrganoP$Result[
                    OrganoP$Analyte == "chlorpyrifos" &
                    OrganoP$Unit == "pg/L"] /(1000*1000)
      # Correct units
       OrganoP$Unit[OrganoP$Analyte == "chlorpyrifos" &
                    OrganoP$Matrix == "water"] <- "ppb"
       
OrganoP %>% filter(Analyte== "chlorpyrifos") %>% distinct(Unit)
```

#### diazinon - ppb

diazinon (2239 records) ppb
diazinon degradate (95 records) ppb
diazinon oxon (14 records) ppb
diazoxon (234 records) ppb

```{r, results = 'hide'}
# diazinon

  # Sediment:	ppb	and ng/g dw	(equivalent units)
    OrganoP$Unit[OrganoP$Analyte == "diazinon" &
                           OrganoP$Matrix == "sediment"] <- "ppb"

  # Water:	pg/L, ng/L, ug/L, ppb
    
      # ppb = pg / (1000*1000)
      # ppb = ng/L / 1000
      # ppb = ug/L
    
      # Convert ng/L to ppb
        OrganoP$Result[OrganoP$Analyte == "diazinon" &
                  OrganoP$Unit == "ng/L"] <- OrganoP$Result[
                    OrganoP$Analyte == "diazinon" &
                    OrganoP$Unit == "ng/L"] /1000
      # Convert pg/L to ppb
        OrganoP$Result[OrganoP$Analyte == "diazinon" &
              OrganoP$Unit == "pg/L"] <- OrganoP$Result[
                OrganoP$Analyte == "diazinon" &
                OrganoP$Unit == "pg/L"] /(1000*1000)
      # Correct units
         OrganoP$Unit[OrganoP$Analyte == "diazinon" &
                      OrganoP$Matrix == "water"] <- "ppb"
         
# Diazinon degradate
    # Correct name
    OrganoP$Analyte[OrganoP$Analyte == "diazinon degradate"] <- "diazinon_degradate"
    unique(OrganoP$Unit[OrganoP$Analyte == "diazinon_degradate"]) 
      
```

```{r, results = 'hide'}
# Diazoxon
    # Convert ng/L to ppb
      OrganoP$Result[OrganoP$Analyte == "diazoxon" &
                OrganoP$Unit == "ng/L"] <- OrganoP$Result[
                  OrganoP$Analyte == "diazoxon" &
                  OrganoP$Unit == "ng/L"] /1000
    # Correct units
      OrganoP$Unit[OrganoP$Analyte == "diazoxon"] <- "ppb"

# Diazinon oxon           
    # Convert ng/L to ppb 
      OrganoP$Result[OrganoP$Analyte == "diazinon oxon" &
                OrganoP$Unit == "ng/L"] <- OrganoP$Result[
                  OrganoP$Analyte == "diazinon oxon" &
                  OrganoP$Unit == "ng/L"] /1000
      
    # Correct units
      OrganoP$Unit[OrganoP$Analyte == "diazinon oxon"] <- "ppb"
    # Correct name
      OrganoP$Analyte[OrganoP$Analyte == "diazinon oxon"] <- "diazinon_oxon"
```
#### malathion - ppb

```{r}
# malathion

  # Sediment:	ppb	and ng/g dw	(equivalent units)
    OrganoP$Unit[OrganoP$Analyte == "malathion" &
                           OrganoP$Matrix == "sediment"] <- "ppb"

  # Water: ng/L, ug/L, ppb
    
      # ppb = ng/L / 1000
      # ppb = ug/L
    
      # Convert ng/L to ppb
        OrganoP$Result[OrganoP$Analyte == "malathion" &
                  OrganoP$Unit == "ng/L"] <- OrganoP$Result[
                    OrganoP$Analyte == "malathion" &
                    OrganoP$Unit == "ng/L"] /1000
      
      # Correct units
         OrganoP$Unit[OrganoP$Analyte == "malathion" &
                      OrganoP$Matrix == "water"] <- "ppb"

OrganoP %>% filter(Analyte== "malathion") %>% distinct(Unit)
```
#### **OrganoP Result**
```{r}
OrganoP %>%
  group_by(Analyte, Matrix, Unit) %>%
  summarise(n = n(),
            mean = mean(Result))
```


```{r}
Wide.OrganoP.Waterdf <- OrganoP %>% filter(!Matrix == "sediment") %>%
  group_by(Date, Latitude, Longitude, Analyte, Matrix) %>%
  summarize(Subregion = first(Subregion),
            Mean = mean(Result, na.rm = T)) %>%
  pivot_wider(names_from = Analyte,
              names_repair = "check_unique",
              values_from = Mean) # Values to fill columns

write.csv(x = Wide.OrganoP.Waterdf , file = "Data/Output/WideSubsets/OrganoP.Wide.water.csv", 
          row.names = F)
```

<br>

### Pyrethroids

```{r}
Pyre <- CEDENSURF %>% filter(SelectList == "Pyrethroids")

Pyre %>%
  group_by(Analyte, Matrix, Unit) %>%
  summarise(n = n(),
            mean = mean(Result))
```

#### cyfluthrin - ppb

```{r, results = 'hide'}
# cyfluthrin

  # Sediment:	ppb	and ng/g dw	(equivalent units)
    Pyre$Unit[Pyre$Analyte == "cyfluthrin" &
                           Pyre$Matrix == "sediment"] <- "ppb"

  # Water: pg/L, ng/L, ug/L, ppb
      # ppb = pg/L /(1000*1000)
      # ppb = ng/L / 1000
      # ppb = ug/L
      
      # Convert pg/L to ppb
        Pyre$Result[Pyre$Analyte == "cyfluthrin" &
              Pyre$Unit == "pg/L"] <- Pyre$Result[
                Pyre$Analyte == "cyfluthrin" &
                Pyre$Unit == "pg/L"] /(1000*1000)
        
      # Convert ng/L to ppb
        Pyre$Result[Pyre$Analyte == "cyfluthrin" &
                  Pyre$Unit == "ng/L"] <- Pyre$Result[
                    Pyre$Analyte == "cyfluthrin" &
                    Pyre$Unit == "ng/L"] /1000
      
      # Correct units
         Pyre$Unit[Pyre$Analyte == "cyfluthrin" &
                      Pyre$Matrix == "water"] <- "ppb"

Pyre %>% filter(Analyte== "cyfluthrin") %>% distinct(Unit)
```

#### esfenvalerate

esfenvalerate
esfenvalerate/fenvalerate
esfenvalerate/fenvalerate-1 (sediment, 4 records)
esfenvalerate/fenvalerate-2 (sediment, 4 records)

```{r, results = 'hide'}
# esfenvalerate

  # Sediment:	ppb	and ng/g dw, and ug/kg	(equivalent units)
    Pyre$Unit[Pyre$Analyte == "esfenvalerate" &
                           Pyre$Matrix == "sediment"] <- "ppb"

  # Water: ng/L, ug/L, ppb
      # ppb = ng/L / 1000
      # ppb = ug/L
        
      # Convert ng/L to ppb
        Pyre$Result[Pyre$Analyte == "esfenvalerate" &
                  Pyre$Unit == "ng/L"] <- Pyre$Result[
                    Pyre$Analyte == "esfenvalerate" &
                    Pyre$Unit == "ng/L"] /1000
      
      # Correct units
         Pyre$Unit[Pyre$Analyte == "esfenvalerate" &
                      Pyre$Matrix == "water"] <- "ppb"

# esfenvalerate/fenvalerate

  # Sediment:	ng/g dw	and	ug/Kg dw (equivalent units)
    Pyre$Unit[Pyre$Analyte == "esfenvalerate/fenvalerate" &
                           Pyre$Matrix == "sediment"] <- "ppb"

  # Water: pg/L, ng/L, ug/L, ppb
      # ppb = pg/L /(1000*1000)
      # ppb = ng/L / 1000
      # ppb = ug/L
      
      # Convert pg/L to ppb
        Pyre$Result[Pyre$Analyte == "esfenvalerate/fenvalerate" &
              Pyre$Unit == "pg/L"] <- Pyre$Result[
                Pyre$Analyte == "esfenvalerate/fenvalerate" &
                Pyre$Unit == "pg/L"] /(1000*1000)
        
      # Convert ng/L to ppb
        Pyre$Result[Pyre$Analyte == "esfenvalerate/fenvalerate" &
                  Pyre$Unit == "ng/L"] <- Pyre$Result[
                    Pyre$Analyte == "esfenvalerate/fenvalerate" &
                    Pyre$Unit == "ng/L"] /1000
      
      # Correct units
         Pyre$Unit[Pyre$Analyte == "esfenvalerate/fenvalerate" &
                      Pyre$Matrix == "water"] <- "ppb"
         
# Correct units for others (sed samples, equivalent to ppb)
  Pyre$Unit[Pyre$Analyte == "esfenvalerate/fenvalerate-1"] <- "ppb"
  Pyre$Unit[Pyre$Analyte == "esfenvalerate/fenvalerate-2"] <- "ppb"

# Correct name
  Pyre$Analyte[Pyre$Analyte == "esfenvalerate/fenvalerate"] <- "esfenvalerate_fenvalerate"
  Pyre$Analyte[Pyre$Analyte == "esfenvalerate/fenvalerate-1"] <- "esfenvalerate_fenvalerate"
  Pyre$Analyte[Pyre$Analyte == "esfenvalerate/fenvalerate-2"] <- "esfenvalerate_fenvalerate"
```

#### **Pyre Result**
```{r}
Pyre %>%
  group_by(Analyte, Matrix, Unit) %>%
  summarise(n = n(),
            mean = mean(Result))
```


```{r}
Wide.Pyre.Waterdf <- Pyre %>% filter(!Matrix == "sediment") %>%
  group_by(Date, Latitude, Longitude, Analyte, Matrix) %>%
  summarize(Subregion = first(Subregion),
            Mean = mean(Result, na.rm = T)) %>%
  pivot_wider(names_from = Analyte,
              names_repair = "check_unique",
              values_from = Mean) # Values to fill columns

write.csv(x = Wide.Pyre.Waterdf , file = "Data/Output/WideSubsets/Pyre.Wide.water.csv", 
          row.names = F)
```
<br>

                     
### GABA inhibitors

(aka anything fipronil?)

```{r, results = 'hide'}

GABA <- CEDENSURF %>% filter(SelectList == "GABA")

GABA %>%
  group_by(Analyte, Matrix, Unit) %>%
  summarise(n = n(),
            mean = mean(Result))
```
#### Constant units, no changes needed

desulfinyl fipronil - ppb
desulfinyl fipronil amide - ppb
fipronil detrifluoromethylsulfinyl - (sediment only) ug/Kg dw


#### fipronil

fipronil - ppb (sed and water)
fipronil amide

```{r, results = 'hide'}
# fipronil

  # Sediment:	ppb	and ng/g dw, and ug/kg	(equivalent units)
    GABA$Unit[GABA$Analyte == "fipronil" &
                           GABA$Matrix == "sediment"] <- "ppb"

  # Water: ng/L, ug/L, ppb
      # ppb = pg/L / (1000*1000)
      # ppb = ng/L / 1000
      # ppb = ug/L
        
      # Convert pg/L to ppb
        GABA$Result[GABA$Analyte == "fipronil" &
              GABA$Unit == "pg/L"] <- GABA$Result[
                GABA$Analyte == "fipronil" &
                GABA$Unit == "pg/L"] /(1000*1000)
        
      # Convert ng/L to ppb
        GABA$Result[GABA$Analyte == "fipronil" &
                  GABA$Unit == "ng/L"] <- GABA$Result[
                    GABA$Analyte == "fipronil" &
                    GABA$Unit == "ng/L"] /1000
      
      # Correct units
         GABA$Unit[GABA$Analyte == "fipronil" &
                      GABA$Matrix == "water"] <- "ppb"
```

fipronil amide - ppb

```{r, results = 'hide'}
# fipronil amide

  # Sediment:	ppb	and ng/g dw, and ug/kg	(equivalent units)
    GABA$Unit[GABA$Analyte == "fipronil amide" &
                           GABA$Matrix == "sediment"] <- "ppb"

  # Water: ug/L and ppb (equivalent units)
    GABA$Unit[GABA$Analyte == "fipronil amide" &
                GABA$Matrix == "water"] <- "ppb"
    
  # Name fix
    GABA$Analyte[GABA$Analyte == "fipronil amide"] <- "fipronil_amide"
```

fipronil desulfinyl & fipronil desulfinyl amide

sediment = ng/g dw
water = ng/L

```{r, results = 'hide'}
# fipronil desulfinyl

  # ng/g dw, and ug/kg	(equivalent units to ppb)
    GABA$Unit[GABA$Analyte == "fipronil desulfinyl" &
                           GABA$Matrix == "sediment"] <- "ng/g dw"

  # Water: pg/L, ng/L, ug/L
      # ng/L = pg/L /1000
      # ng/L= ug/L  *1000
        
      # Convert pg/L to ng/L
        GABA$Result[GABA$Analyte == "fipronil desulfinyl" &
              GABA$Unit == "pg/L"] <- GABA$Result[
                GABA$Analyte == "fipronil desulfinyl" &
                GABA$Unit == "pg/L"] /1000
        
      # Convert ug/L to ng/L
        GABA$Result[GABA$Analyte == "fipronil desulfinyl" &
                  GABA$Unit == "ug/L"] <- GABA$Result[
                    GABA$Analyte == "fipronil desulfinyl" &
                    GABA$Unit == "ug/L"]*1000
      
      # Correct units
         GABA$Unit[GABA$Analyte == "fipronil desulfinyl" &
                      GABA$Matrix == "water"] <- "ng/L"
        
# fipronil desulfinyl amide

      # Convert ug/L to ng/L
        GABA$Result[GABA$Analyte == "fipronil desulfinyl amide" &
                  GABA$Unit == "ug/L"] <- GABA$Result[
                    GABA$Analyte == "fipronil desulfinyl amide" &
                    GABA$Unit == "ug/L"]*1000
      
      # Correct units
         GABA$Unit[GABA$Analyte == "fipronil desulfinyl amide" &
                      GABA$Matrix == "water"] <- "ng/L"
         
```

#### fipronil sulfide - ppb

```{r, results = 'hide'}
# fipronil sulfide

  # Sediment:	ppb	and ng/g dw, and ug/kg	(equivalent units)
    GABA$Unit[GABA$Analyte == "fipronil sulfide" &
                           GABA$Matrix == "sediment"] <- "ppb"

  # Water: ng/L, ug/L, ppb
      # ppb = pg/L / (1000*1000)
      # ppb = ng/L / 1000
      # ppb = ug/L
        
      # Convert pg/L to ppb
        GABA$Result[GABA$Analyte == "fipronil sulfide" &
              GABA$Unit == "pg/L"] <- GABA$Result[
                GABA$Analyte == "fipronil sulfide" &
                GABA$Unit == "pg/L"] /(1000*1000)
        
      # Convert ng/L to ppb
        GABA$Result[GABA$Analyte == "fipronil sulfide" &
                  GABA$Unit == "ng/L"] <- GABA$Result[
                    GABA$Analyte == "fipronil sulfide" &
                    GABA$Unit == "ng/L"] /1000
      
      # Correct units
         GABA$Unit[GABA$Analyte == "fipronil sulfide" &
                      GABA$Matrix == "water"] <- "ppb"
        
```

fipronil sulfone

```{r, results = 'hide'}
# fipronil sulfone

  # Sediment:	ppb	and ng/g dw, and ug/kg	(equivalent units)
    GABA$Unit[GABA$Analyte == "fipronil sulfone" &
                           GABA$Matrix == "sediment"] <- "ppb"

  # Water: ng/L, ug/L, ppb
      # ppb = pg/L / (1000*1000)
      # ppb = ng/L / 1000
      # ppb = ug/L
        
      # Convert pg/L to ppb
        GABA$Result[GABA$Analyte == "fipronil sulfone" &
              GABA$Unit == "pg/L"] <- GABA$Result[
                GABA$Analyte == "fipronil sulfone" &
                GABA$Unit == "pg/L"] /(1000*1000)
        
      # Convert ng/L to ppb
        GABA$Result[GABA$Analyte == "fipronil sulfone" &
                  GABA$Unit == "ng/L"] <- GABA$Result[
                    GABA$Analyte == "fipronil sulfone" &
                    GABA$Unit == "ng/L"] /1000
      
      # Correct units
         GABA$Unit[GABA$Analyte == "fipronil sulfone" &
                      GABA$Matrix == "water"] <- "ppb"
         
# Correct names

  GABA$Analyte[GABA$Analyte == "desulfinyl fipronil"] <- "fipronil_desulfinyl"
  GABA$Analyte[GABA$Analyte == "desulfinyl fipronil amide"] <- 
    "fipronil_desulfinyl_amide"
    
  GABA$Analyte[GABA$Analyte == "fipronil desulfinyl"] <- "fipronil_desulfinyl"
  GABA$Analyte[GABA$Analyte == "fipronil desulfinyl amide"] <-
    "fipronil_desulfinyl_amide"
  GABA$Analyte[GABA$Analyte == "fipronil sulfide"] <- "fipronil_sulfide"
  GABA$Analyte[GABA$Analyte == "fipronil sulfone"] <- "fipronil_sulfone"

# Remove if insufficient replication
  GABA <- GABA %>% filter(!Analyte == "fipronil detrifluoromethylsulfinyl")
```

#### **GABA Result**
```{r}
GABA %>%
  group_by(Analyte, Matrix, Unit) %>%
  summarise(n = n(),
            mean = mean(Result))
```


```{r}
Wide.GABA.Waterdf <- GABA %>% filter(!Matrix == "sediment") %>%
  group_by(Date, Latitude, Longitude, Analyte, Matrix) %>%
  summarize(Subregion = first(Subregion),
            Mean = mean(Result, na.rm = T)) %>%
  pivot_wider(names_from = Analyte,
              names_repair = "check_unique",
              values_from = Mean) # Values to fill columns

write.csv(x = Wide.GABA.Waterdf , file = "Data/Output/WideSubsets/GABA.Wide.water.csv", 
          row.names = F)
```

<br>

### Others: Glyphosate, Atrazine, Neonicitinoids

```{r}
Other <- CEDENSURF %>% filter(SelectList == c("Glyphosate", "Atrazine", "Neon"))
```

#### Neonicotinoids 

All measures are in water. 

hydroxy-imidacloprid (2 records total)
imidacloprid (228 records)

```{r, results = 'hide'}
Other %>% filter(SelectList == "Neon") %>%
  group_by(Analyte, Matrix, Unit) %>%
  summarise(n = n(),
            mean = mean(Result))
```

```{r, results = 'hide'}
  # Water: ng/L, ug/L, ppb

      # ppb = ng/L / 1000
      # ppb = ug/L

      # Convert ng/L to ppb
        Other$Result[Other$SelectList == "Neon" &
                  Other$Unit == "ng/L"] <- Other$Result[
                    Other$SelectList == "Neon" &
                    Other$Unit == "ng/L"] /1000
      
      # Correct units
         Other$Unit[Other$SelectList == "Neon"] <- "ppb"
         
      # Remove records with insufficient replication
         Other <- Other %>% filter(!Analyte == "hydroxy-imidacloprid")
    
```

#### Glyphosate - ppb

(water only) 219 records, no conversion needed (equivalent units)

```{r, results = 'hide'}
Other %>% filter(SelectList == "Glyphosate") %>%
  group_by(Analyte, Matrix, Unit) %>%
  summarise(n = n(),
            mean = mean(Result))

# units ug/L and ppb (equivalent units)
  Other$Unit[Other$SelectList == "Glyphosate"] <- "ppb"
```

#### Atrazine

Atrazine (sed and water) - ppb
Atrazine degradate (water only) - ppb

```{r, results = 'hide'}
Other %>% filter(SelectList == "Atrazine") %>%
  group_by(Analyte, Matrix, Unit) %>%
  summarise(n = n(),
            mean = mean(Result))
```
```{r, results = 'hide'}
# Water: ng/L, ug/L, ppb

      # ppb = ng/L / 1000
      # ppb = ug/L

      # Convert ng/L to ppb
        Other$Result[Other$SelectList == "Atrazine" &
                  Other$Unit == "ng/L"] <- Other$Result[
                    Other$SelectList == "Atrazine" &
                    Other$Unit == "ng/L"] /1000
      
      # Correct units
         Other$Unit[Other$SelectList == "Atrazine"] <- "ppb"
         
      # Correct name
        Other$Analyte[Other$Analyte == "atrazine degradate"] <- "atrazine_degradate"
```

#### **Other Result**
```{r}
Other %>%
  group_by(Analyte, Matrix, Unit) %>%
  summarise(n = n(),
            mean = mean(Result))
```


```{r}
Wide.Other.Waterdf <- Other %>% filter(!Matrix == "sediment") %>%
  group_by(Date, Latitude, Longitude, Analyte, Matrix) %>%
  summarize(Subregion = first(Subregion),
            Mean = mean(Result, na.rm = T)) %>%
  pivot_wider(names_from = Analyte,
              names_repair = "check_unique",
              values_from = Mean) # Values to fill columns

write.csv(x = Wide.Other.Waterdf , file = "Data/Output/WideSubsets/Other.Wide.water.csv", 
          row.names = F)
```

<br>


### Save

Compile and Save Modified Dataset (Long)

The file Data/Output/CEDENSURF_Limited_FixedUnits.csv contains all corrected units and analytes chosen for our model in long format. 

```{r}
Limited <- rbind(WQP, GABA, Metal, OrganoP, Pyre, Other)

write.csv(x = Limited, 
          file = "Data/Output/CEDENSURF_Limited_FixedUnits.csv", 
          na = "", row.names = F)
```

## 3. Summarize into wide format

I used pivot_wider to summarize analyte results by date and subregion within each matrix. 

Because column names are the means by which Netica differentiates nodes, I made the wide format data from the water-matrix contain just the analyte as the column name, while the sediment data includes analyte_sediment as the column names

Allwater.Wide.csv contains the final summarized analytes in wide format (one row per date/location; one column per analyte)

Allsed.Wide.csv contains the final summarized analytes in wide format (one row per date/location; one column per analyte)

```{r}
## Subset: Water Matrix

# Convert long to wide
Wide.waterdf <- Limited %>% filter(!Matrix == "sediment") %>%
  group_by(Date, Latitude, Longitude, Analyte, Matrix) %>%
  summarize(Subregion = first(Subregion),
            Mean = mean(Result, na.rm = T)) %>%
  pivot_wider(names_from = Analyte,
              names_repair = "check_unique",
              values_from = Mean) # Values to fill columns

# save document
write.csv(x = Wide.waterdf, file = "Data/Output/Allwater.Wide.csv", 
          row.names = F)
```

```{r}
## Subset: Water Matrix

# Convert long to wide
Wide.seddf <- Limited %>% filter(!Matrix == "water") %>%
  group_by(Date, Latitude, Longitude, Analyte, Matrix) %>%
  summarize(Subregion = first(Subregion),
            Mean = mean(Result, na.rm = T)) %>%
  pivot_wider(names_from = c(Analyte,Matrix),
              names_repair = "check_unique",
              values_from = Mean) # Values to fill columns

# save document
write.csv(x = Wide.seddf, file = "Data/Output/Allsed.Wide.csv", 
          row.names = F)
```
