---
title: "Chemical Stressors - Toxicological Units"
author: "Erika W"
date: "5/1/2021"
output:
  html_document:
    code_download: true
    keep_md: true
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=F, message=F}
rm(list = ls())
library(data.table)
library(lubridate)
library(sf)
library(tidyverse)
```

## Intro

This rmd defines the toxicological unit (Molar Mass) for chemicals that may need to be combined later. 

## Load Data

```{r}
# Load CEDENSURF Limited Data
AnalyteList <- fread("https://github.com/WWU-IETC-R-Collab/CEDENSURF-mod/raw/main/Data/Output/CEDENSURF_Limited_FixedUnits.csv") %>% 
  distinct(Analyte, Unit, Matrix, SelectList)
```

## Molar Mass

We also may need to append toxicological units onto the data, in order to combine multiple analytes falling under the same category. 

If we're going to ADD toxicological units, I don't believe we should use analytes that were only measured in one region (ie, would falsly inflate that regions relative risk just by having more information), unless we were to estimate them for other regions by proxy (ie: land use)

diazinon MM = 304.34 g/mol
diazoxon MM = 288.28 g/mol

esfenvalerate MM = 419.91

fipronil = 437.1 g/mol

fipronil amide = 455.2 g/mol
fipronil desulfinyl MM = 389.08 g/Mol
fipronil desulfinyl amide MM =  ???

fipronil sulfone MM = 453.1 g/mol
fipronil sulfide MM = 421.1 g/mol

```{r}
AnalyteList$MolarMass <- NA

# Organophosphates

AnalyteList$MolarMass[AnalyteList$Analyte == "diazinon"] = 304.34

AnalyteList$MolarMass[AnalyteList$Analyte == "diazoxon"] = 288.28

# Pyrethroids

AnalyteList$MolarMass[AnalyteList$Analyte == "esfenvalerate"] = 419.91

# GABA Inhibitors

AnalyteList$MolarMass[AnalyteList$Analyte == "fipronil"] = 437.1

AnalyteList$MolarMass[AnalyteList$Analyte == "fipronil_amide"] = 455.2

AnalyteList$MolarMass[AnalyteList$Analyte == "fipronil_desulfinyl"] = 389.08

AnalyteList$MolarMass[AnalyteList$Analyte == "fipronil_sulfone"] = 453.1

AnalyteList$MolarMass[AnalyteList$Analyte == "fipronil_sulfide"] = 421.1

AnalyteList$MolarMass[AnalyteList$Analyte == ""] = 
  
AnalyteList$MolarMass[AnalyteList$Analyte == ""] = 
  
AnalyteList$MolarMass[AnalyteList$Analyte == ""] = 
```

