---
title: "Merge Investigation"
author: "Erika W"
date: "2/18/2021"
output:
  html_document:
    code_download: true
    keep_md: true
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=F, message=F}
rm(list = ls())
library(data.table)
library(lubridate)
library(sf)
library(tidyverse)
```

# Load & Prep Data

### CEDEN

Two files - one with tox data, and one with wq data

```{r}
# Load CEDEN Data
CEDENMod_Tox <- fread("https://github.com/WWU-IETC-R-Collab/CEDEN-mod/raw/main/Data/Output/CEDENMod_Toxicity.csv")

CEDENMod_WQ <- fread("https://github.com/WWU-IETC-R-Collab/CEDEN-mod/raw/main/Data/Output/CEDENMod_WQ.csv")
```

### SURF

Two files - one with wq data, and one with SED data ?
```{r}
SURFMod_SED <- fread("https://github.com/WWU-IETC-R-Collab/CEDENSURF-mod/raw/main/Data/Output/SURFMod_SED.csv")

SURFMod_WQ <- fread("https://github.com/WWU-IETC-R-Collab/CEDENSURF-mod/raw/main/Data/Output/SURFMod_water.csv")
```

### Append with source

```{r}
CEDENMod_Tox$Source <- rep("CEDEN", times=nrow(CEDENMod_Tox))

CEDENMod_WQ$Source <- rep("CEDEN", times=nrow(CEDENMod_WQ))

SURFMod_SED$Source <- rep("SURF", times=nrow(SURFMod_SED))

SURFMod_WQ$Source <- rep("SURF", times=nrow(SURFMod_WQ))
```

# Remove Duplicates

Due to reported combined efforts to translate CEDEN data to SURF and vice versa, and issues with replicates being retained in each dataset, careful detection and elimination of duplicates should precede any merge and analysis.

## CEDEN Data

The tox data seems to be focused on associating WQ parameters with specific biological assays. These WQ parameters may already be present in the CEDEN_WQ dataset. 

There are `r nrow(CEDENMod_WQ)` records in the WQ dataset
and `r nrow(CEDENMod_Tox)` in the Tox dataset. 

### CEDEN Tox 

In the tox dataset, there were several records related to each species assessment - by creating a group identifier by the location, date, and organism assessed, I calculated that there were ~ 2430 unique biological assays.`r length(unique(CEDENMod_Tox$OrganismName))` different species are represented.

#### Exact duplicates

duplicated() produces a vector identifying which records are duplicates. This can be appended to the original dataset for further investigation. The first record is marked FALSE, while subsequent matches are marked TRUE

```{r}
# add index column identifying which entries are duplicated

CEDENMod_Tox$DupCheck <- duplicated(CEDENMod_Tox)

summary(CEDENMod_Tox$DupCheck)
```

Is it only some programs that have duplication? 9/10 programs do.

```{r}
length(unique(CEDENMod_Tox$Program)) # 10

unique(CEDENMod_Tox$Program[CEDENMod_Tox$DupCheck == "TRUE"])
```

#### Alternative (looser) assessment of duplication

If we assume that some columns may differ due to differences in data loading, then we may want to use a looser structure to detect duplicates. 

By assuming that records in the same location on the same date,  measuring the same analyte via the same collection method, and obtaining the same result are duplicates, we find almost 50% more duplicates. 
25729 records, to be specific. 

That is `r 100*25729/nrow(CEDENMod_Tox)`% of the dataset.

```{r}
NoDup_Tox<- distinct(CEDENMod_Tox, Date, StationName, Analyte, CollectionMethod, Result, .keep_all= TRUE)

nrow(CEDENMod_Tox) - nrow(NoDup_Tox)
```

Since we are using the df for the water parameters and not the associated organism survival, we can remove the organism column, and also remove records that assess biotic parameters like 'survival' and 'biomass'.

```{r}
NoDup_Tox <- NoDup_Tox %>% filter(Analyte != "Survival") %>%
  filter(Analyte != "Biomass (wt/orig indiv)") %>%
  filter(Analyte != "Young/female") %>%
  filter(Analyte != "Total Cell Count") %>%
  select(-OrganismName)
```

We're left with only `r nrow(NoDup_Tox)` unique, useful records in the tox dataset - or `r round(c(100*nrow(NoDup_Tox)/nrow(CEDENMod_Tox)),1)`% of the original tox data remaining

<br>

### CEDEN WQ

#### Exact duplicates

duplicated() produces a vector identifying which records are duplicated. This can be appended to the original dataset for further investigation.

```{r}
# add index column identifying which entries are duplicated
CEDENMod_WQ$DupCheck <- duplicated(CEDENMod_WQ)

summary(CEDENMod_WQ$DupCheck)
```

This method identified only 657 exact duplicates in the entire WQ dataset.

Only 9/17 programs present these exact duplicates

```{r}
length(unique(CEDENMod_WQ$Program)) # 17
unique(CEDENMod_WQ$Program[CEDENMod_WQ$DupCheck == "TRUE"])
```

#### Looser assessment of Duplication

Utilizing the distinct() function to assume that records in the same location on the same date, measuring the same analyte via the same collection method and obtaining the same result are duplicates, we find 1536 duplicate records.

That is more than double the number of exact duplicates found, yet still only `r round(c(100*1536/nrow(CEDENMod_WQ)),1)`% of the entire WQ dataset.

```{r}
# Remove duplicate rows of the dataframe using multiple variables

NoDupWQ <- distinct(CEDENMod_WQ, Date, Analyte, StationName, CollectionMethod, Result, .keep_all= TRUE)

nrow(CEDENMod_WQ) - nrow(NoDupWQ)
```
<br>

<br>

## Merging CEDEN data

After dealing with duplication WITHIN the CEDEN tox and wq datasets, there were only 9 duplicate records found following the merged data. (75 if Collection Method is not a requirement for establishing duplication)

```{r}
# Strip column with duplicate check from both, and organism from tox
NoDup_Tox <- select(NoDup_Tox, -c(DupCheck))
NoDupWQ <- select(NoDupWQ, -DupCheck)

# Vector of column names to compare
WQ <- names(NoDupWQ)
TOX <- names(NoDup_Tox)

#Add missing columns to CEDEN WQ
DIF<- setdiff(TOX, WQ) # gives items in T that are not in W
NoDupWQ[, DIF] <- NA

#Add missing columns to CEDEN TOX
DIF<- setdiff(WQ, TOX) # gives items in W that are not in T
NoDup_Tox[, DIF] <- NA

# Finishing touches before merge; order columns to match - is this really necessary for merge? IDK

WQ <- sort(names(NoDupWQ))
TOX <- sort(names(NoDup_Tox))

NoDup_Tox <- NoDup_Tox %>% select(TOX)
NoDupWQ <- NoDupWQ %>% select(WQ)

# Check once all columns have perfect matches?
# tibble(SURF = names(NoDup_Tox), CEDEN = names(NoDupWQ))

# MERGE
CEDEN_ALL <- rbind(NoDupWQ,NoDup_Tox)
```

```{r}
# Ideas from: https://www.datasciencemadesimple.com/remove-duplicate-rows-r-using-dplyr-distinct-function/

# Remove duplicate rows of the dataframe using multiple variables

CEDEN_ALL_DupChecked <- distinct(CEDEN_ALL, Date, Analyte, CollectionMethod, StationName, Result, .keep_all= TRUE)

nrow(CEDEN_ALL)-nrow(CEDEN_ALL_DupChecked)
```
### Duplication Errors? {.tabset}

#### Queries

Are there duplicates retained by restricting to unique Collection Methods, which should not be?

There are 401 extra records retained when using Collection method to differentiate identical results. 343 of these duplicate records were not zero-result values, and therefore represent records of concern.

```{r}
CEDEN_ALL_DupChecked <- distinct(CEDEN_ALL, Date, Analyte, 
                                 CollectionMethod, StationName, 
                                 Result, .keep_all= TRUE)

CEDEN_ALL_Check1 <- distinct(CEDEN_ALL, Date, Analyte, 
                              StationName, Result, 
                             .keep_all= TRUE)

nrow(CEDEN_ALL_DupChecked) - nrow(CEDEN_ALL_Check1)
```
Restricting the duplicates of concern to those which are non-zero values... It appears that there are 6 different collection methods represented in the duplicated data.

Sleuthing what is going on:

1. Many are duplication within a parent project, with one entry having sample method marked "not recorded" and the other defined. While the entire CEDEN dataset has 305 samples labeled "not recorded", 153 of these are likely duplicated data.

**Question** Is it worth simply removing all records for which the sample method is not recorded?

2. Two projects hold the majority of the duplication: CA Dept of Transit NPDES Monitoring 2014-2015 (153), and San Joaquin County Delta Water Quality Coalition (105)

3. Duplication within the Regional Monitoring Program - Status and Trends deserves closer investigation. Those samples have identical, >0 results for samples labeled with:

+ Collection_Method = Sediment_Core & Matrix = sediment
+ Collection_Method = Sediment_Grab & Matrix = overlyingwater

**--> Question** Does it make sense to have a sample defined as a sediment grab method + matrix of overlying water? Does it make sense for these samples to have identical results?

```{r}
# Find columns retained only when restricting Collection Method also

DIF<- setdiff(CEDEN_ALL_DupChecked, CEDEN_ALL_Check1) # gives items in A that are not in B


# Restrict to non-zero results (concern for duplication)

DIF <- DIF %>% filter(., Result != "0") 
nrow(DIF)

# Summary of collection methods which differentiate these duplicates. 

DIF$CollectionMethod <- as.factor(DIF$CollectionMethod)
summary(DIF$CollectionMethod)

# Summary of parent projects represented in the duplicated data

DIF$ParentProject <- as.factor(DIF$ParentProject)
summary(DIF$ParentProject)
```

#### Printout of duplicated data

Again, these duplicates are retained in the current methodology due to differences between Collection Method. The question is whether they are truly different records, or whether we need to further refine the selection criteria.

```{r}
# Bring back original data associated with each duplicate identified

Dups <- vector("list") # save empty list where each subset be saved

for (i in 1:343){
  Dups[[i]] <- CEDEN_ALL %>% filter(., Analyte == DIF$Analyte[i],
                                         Date == DIF$Date[i], 
                                         StationName == DIF$StationName[i],
                                    Result == DIF$Result[i])
  }

Example <- do.call(rbind, Dups)
Example

# How does number of "not recorded" methods in these duplications compare to those in the entire df?

paste("Collection Method = Not Recorded in...")

tibble(Entire = nrow(CEDEN_ALL_DupChecked[CEDEN_ALL_DupChecked$CollectionMethod== "Not Recorded"]), Dup = nrow(Example[Example$CollectionMethod== "Not Recorded"]))
```

### Fix CEDEN nomenclature

**Split Analyte Column**

Because of formatting differences between the amount of data recorded under "Analyte" in CEDEN compared to the "Chemical_name" in SURF (which will be renamed Analyte), we opted to split the data in CEDEN's analyte column into two columns: 

Analyte (Chemical Name), and Analyte_Type (ie: total or particulate)

Using separate() to split the column and requiring the separation to be a comma and space ", " seems to have worked well, except for one name which appears to be empty

**Convert to lower-case**

SURF chemicals are all lowercase. We can make all letters in the CEDEN data lowercase using tolower() so that they will be compatible.

```{r}
# Remove subset of data determined to be >50% duplication

CEDEN_ALL_DupChecked <- CEDEN_ALL_DupChecked %>% filter(CollectionMethod != "Not Recorded")

# Split Analyte column

CEDEN_ALL_DupChecked <- CEDEN_ALL_DupChecked %>%
  separate(Analyte, into = c("Analyte", "Analyte_type"), sep = ", " , extra = "merge")

# Convert to lowercase

CEDEN_ALL_DupChecked$Analyte <- tolower(CEDEN_ALL_DupChecked$Analyte)

# Preview
head(sort(unique(CEDEN_ALL_DupChecked$Analyte))) # 908 unique Analytes total

# Looks like requiring the separation from extra to analyte to contain a comma and a space allowed the full names to be retained. Without that, the separation led to an analyte "1" which should have been 1,2-bis(2,4,6- tribromophenoxy)ethane, etc.
```

This simplification of Analyte name would lead to 15,290 more records being considered "duplication" using our current method, mostly (14,776) containing a zero-result. Because these differences in analytes are not retained in the SURF dataset, it makes sense to condense them (remove duplicates) prior to merging the databases.

```{r, eval = F}
Check <- distinct(CEDEN_ALL_DupChecked, Date, Analyte, CollectionMethod, StationName, Result, .keep_all= TRUE)

nrow(CEDEN_ALL_DupChecked) - nrow(Check)

DIF<- setdiff(CEDEN_ALL_DupChecked, Check)
length(DIF$Result[DIF$Result == "0"])
```

Removal of these simplified duplicates also eliminates the utility of retaining the "Analyte Type" column. For example, a reading of Analyte X with Type = Total Type = Suspended have the exact same result. Removing duplicates would keep the first record (Analyte = X, Type = Total) and remove the second (Analyte X, Type = S). In the dataframe, if you are trying to reason backwards to the meaning of that remaining meaning (Analyte X, Type = Total), you're missing the other half of the story (Type = Sus too). So, to avoid improper interpretation of this dataframe, Analyte Type should be removed. 

```{r}
CEDEN_ALL_DupChecked <- distinct(CEDEN_ALL_DupChecked, Date, Analyte, CollectionMethod, StationName, Result, .keep_all= TRUE) %>%
  select(-Analyte_type)

# 127,874
```

### CEDEN merge result

Using these QA/QC methods, `r nrow(CEDEN_ALL_DupChecked)` unique records are available through the CEDEN datasets. 

<br>

<br>


## SURF data {.tabset}

There are `r nrow(SURFMod_WQ)` records in the WQ dataset
and `r nrow(SURFMod_SED)` in the SED dataset. 

### SURF Water

#### Data Prep

**Correct mismatched column names**

Rather than using Steven's method of renaming using a list (which is dependent on an expected number and order of column names), I renamed each column individually

```{r}
SURFMod_WQ$Unit <- "ppb"

SURFMod_WQ <- SURFMod_WQ %>% rename(Date = Sample_date,
          Analyte = Chemical_name, 
          Result = Concentration..ppb., 
          CollectionMethod = Sample_type, 
          StationCode = Site_code,
          StationName = Site_name,
          MDL = Method_detection_level..ppb.,
          LOQ = Level_of_quantification..ppb.)
```

#### Exact duplicates

duplicated() produces a vector identifying which records are duplicated. This can be appended to the original dataset for further investigation.

```{r}
# add index column identifying which entries are duplicated

SURFMod_WQ$DupCheck <- duplicated(SURFMod_WQ)

summary(SURFMod_WQ$DupCheck)
```

Fantastic. There are no exact duplicates within the SURF water data - even when we include the CEDEN allowed in SURF

#### Looser assessment of Duplication

Utilizing the distinct() function to assume that records in the same location on the same date, measuring the same analyte and obtaining the same result are duplicates, we find 17756 (there were only 5,676 w/o CEDEN data) duplicate records.

```{r}
# Remove duplicate rows of the dataframe using multiple variables

NoDupWQ <- distinct(SURFMod_WQ, Date, Analyte, CollectionMethod, StationName, Result, .keep_all= TRUE)

nrow(SURFMod_WQ) - nrow(NoDupWQ)
```
<br>

<br>

### SURF sediment

#### Data Prep

**Correct mismatched column names**
```{r}
SURFMod_SED$Unit <- "ppb"

SURFMod_SED <- SURFMod_SED %>% rename(Date = Sample_date,
          Analyte = Chemical_name, 
          Result = Concentration..ppb., 
          CollectionMethod = Sample_type, 
          StationCode = Site_code,
          StationName = Site_name,
          MDL = Method_detection_level..ppb.,
          LOQ = Level_of_quantification..ppb.)
```

#### Exact duplicates

duplicated() produces a vector identifying which records are duplicated. This can be appended to the original dataset for further investigation.

```{r}
# add index column identifying which entries are duplicated

SURFMod_SED$DupCheck <- duplicated(SURFMod_SED)

summary(SURFMod_SED$DupCheck)
```

Fantastic. There are no exact duplicates within the SURF sediment data, even with CEDEN also contained.

#### Looser assessment of Duplication

Utilizing the distinct() function to assume that records in the same location on the same date, measuring the same analyte and obtaining the same result are duplicates, we find only 4,120 (was 784 without CEDEN sources)

```{r}
# Remove duplicate rows of the dataframe using multiple variables

NoDupSED <- distinct(SURFMod_SED, Date, Analyte, CollectionMethod, StationName, Result, .keep_all= TRUE)

nrow(SURFMod_SED) - nrow(NoDupSED)
```
<br>

<br>

## Merge SURF df

```{r}
# Strip column with duplicate check from both
NoDupSED <- select(NoDupSED, -DupCheck)
NoDupWQ <- select(NoDupWQ, -DupCheck)

# Vector of column names to compare
WQ <- names(NoDupWQ)
SED <- names(NoDupSED)

#Add missing columns to CEDEN WQ
DIF<- setdiff(SED, WQ) # gives items in S that are not in W
NoDupWQ[, DIF] <- NA

#Add missing columns to CEDEN SED
DIF<- setdiff(WQ, SED) # gives items in W that are not in S
NoDupSED[, DIF] <- NA

# Finishing touches before merge; order columns to match - is this really necessary for merge? IDK

WQ <- sort(names(NoDupWQ))
SED <- sort(names(NoDupSED))

NoDupSED <- NoDupSED %>% select(SED)
NoDupWQ <- NoDupWQ %>% select(WQ)

# Check once all columns have perfect matches?
# tibble(SURF = names(NoDupSED), CEDEN = names(NoDupWQ))

# MERGE
SURF_ALL <- rbind(NoDupWQ,NoDupSED)
```

### Duplication between SURF sets

ZERO duplication found between the SED and WQ datasets, assuming duplicates would have to have the exact same Location, Date, Analyte, Collection Method, and Result.

```{r}
# Ideas from: https://www.datasciencemadesimple.com/remove-duplicate-rows-r-using-dplyr-distinct-function/

# Remove duplicate rows of the dataframe using multiple variables

SURF_ALL_DupChecked <- distinct(SURF_ALL, Date, Analyte, CollectionMethod, StationName, Result, .keep_all= TRUE)

nrow(SURF_ALL)-nrow(SURF_ALL_DupChecked)
```

### Request: Examples of SURF duplicates when Collection Method is omitted? {.tabset}

#### With CEDEN Data

Able to make df of "duplicate" entries for further inspection. Though **ALL WERE DUE TO RESULT == 0 WITHOUT CEDEN INCLUDED**

When retaining data sourced from CEDEN, there are some instances of results != 0 with identical results.

```{r}
# Create DF limited to only one record per date,loc,analyte,result combo

SURF_ALL_Check1 <- distinct(SURF_ALL, Date, Analyte, StationName, Result, .keep_all= TRUE)

# ID columns retained only when use Collection Method also

DIF<- setdiff(SURF_ALL_DupChecked, SURF_ALL_Check1) # gives items in A that are not in B

# Produce vector of all different "Result" values in the duplicated data

length(unique(SURF_ALL$Result)) # how many different "result" values are represented in the entire combined SURF dataset

unique(DIF$Result) # All unique "result" values within the duplicates that were removed.
```

These duplicates have some commonalities;

1. they are all brought into SURF from "CEDEN" (except for one "OTHER")

2. they have different "collection methods", typically one with a collection method specified (Study CD 523), and one without the method specified (Study CD 305)

3. Duplicates are almost all between Study CD's 523 & 305. Only one pair of replicates were due to Study CD's 451 & 80. 

```{r}

DIF <- DIF %>% filter(., Result != "0") # Only 19 duplicates were removed that had non-0-result values

# Bring back original data associated with each duplicate identified

Dups <- vector("list") # save empty list where each subset be saved

for (i in 1:19){
  Dups[[i]] <- SURF_ALL %>% filter(., Analyte == DIF$Analyte[i],
                                         Date == DIF$Date[i], 
                                         StationName == DIF$StationName[i])
  }

Example <- do.call(rbind, Dups)
Example
```

```{r}
unique(Example$Data.source)

unique(Example$Study_cd)
```

Given this info, it is worthwhile to ask if there are any unique entries by Study_cd 305, or if all are duplications of more complete data by Study_523.

**ALL ENTRIES IN SURF BY STUDY CD 305 ARE DUPS**

```{r}

length(SURF_ALL$Study_cd[SURF_ALL$Study_cd == "305"]) # Number of entries by this Agency/Program in all SURF data

length(Example$Study_cd[SURF_ALL$Study_cd == "305"]) # Number of entries that were part of the duplication findings
```

#### Without CEDEN 

ZERO duplication found between the SED and WQ datasets, assuming duplicates would have to have the exact same Location, Date, Analyte, Collection Method, and Result.

```{r}
# Ideas from: https://www.datasciencemadesimple.com/remove-duplicate-rows-r-using-dplyr-distinct-function/

# Filter out data sourced from CEDEN

SURF_ALL_NC <- filter(SURF_ALL, Data.source != "CEDEN")
# Remove duplicate rows of the dataframe using multiple variables

SURF_ALL_DupCheckedNC <- distinct(SURF_ALL_NC, Date, Analyte, CollectionMethod, StationName, Result, .keep_all= TRUE)

nrow(SURF_ALL_NC)-nrow(SURF_ALL_DupCheckedNC)
```

*Note* I had started without 'Collection Method' in the distinct() query, but added it as an additional requirement for establishing duplication because without that record, the difference between results for water vs sediment were not recognized and **26127** records were considered duplicates following the merging of the SURF data.

This is still surprising to me, because that means the RESULT concentration was identical for both sediment and water samples. 

By creating a df of "duplicate" entries for further inspection. Findings confirm that
**ALL ARE DUE TO RESULT == 0**

```{r}
# Create DF limited to only one record per date,loc,analyte,result combo

SURF_ALL_Check1 <- distinct(SURF_ALL_NC, Date, Analyte, StationName, Result, .keep_all= TRUE)

# ID columns retained only when use Collection Method also

DIF<- setdiff(SURF_ALL_DupCheckedNC, SURF_ALL_Check1) # gives items in A that are not in B

# Produce vector of all different "Result" values in the duplicated data

length(unique(SURF_ALL_NC$Result)) # how many different "result" values are represented in the entire combined SURF dataset

unique(DIF$Result) # All unique "result" values within the duplicates that were removed.
```

### Result of SURF Merge

Due to duplication found above, remove study 305 that was entirely duplication prior to next steps (Because this duplication was not able to be removed by simply filtering on date, loc, analyte, collection method, and result)

```{r}
SURF_ALL_DupChecked <- filter(SURF_ALL_DupChecked, Study_cd != "305")

nrow(SURF_ALL_DupChecked)
```
There are 143,000 unique records available through SURF.

<br>

<br>

# Merge SURF and CEDEN
Now that each dataset has been independently inspected for duplication *within* each dataset, they can be merged and searched for duplication *between* the datasets.

### Data prep

**Match columns to CEDEN data**

```{r}
C <- names(CEDEN_ALL_DupChecked)
S <- names(SURF_ALL_DupChecked)

DIF<- setdiff(S, C) # gives items in S that are not in C
```

Columns that were in SURF and not CEDEN:
`r DIF`

```{r}
#Add missing columns to CEDEN
CEDEN_ALL_DupChecked[, DIF] <- NA
```

```{r}
#Add missing columns to SURF
DIF<- setdiff(C, S) # gives items in C that are not in S
SURF_ALL_DupChecked[, DIF] <- NA
```

Columns that were in CEDEN and not SURF:
` r DIF`

### Merge

Reorder columns to align, then merge with rbind()

```{r}
# Finishing touches before merge; order columns to match - is this really necessary for merge? IDK

C <- sort(names(CEDEN_ALL_DupChecked)) # 908
S <- sort(names(SURF_ALL_DupChecked)) # 327

SURF_ALL_DupChecked <- SURF_ALL_DupChecked %>% select(S)

CEDEN_ALL_DupChecked <- CEDEN_ALL_DupChecked %>% select(C)

# Check once all columns have perfect matches?
# tibble(SURF = names(SURF_ALL_DupChecked), CEDEN = names(CEDEN_ALL_DupChecked))

# MERGE
CEDENSURF <- rbind(SURF_ALL_DupChecked, CEDEN_ALL_DupChecked)
```

### Check for duplicates {.tabset}

Retaining the 45,064 records of CEDEN data in SURF prior to merge, this method removes 15,290.

This means `r 45064 - 15290` records of SURF's CEDEN-sourced data are retained. Some may be unique, and some may be additional duplicates that we are not catching. 

```{r}
# How many SURF records are sourced from CEDEN?
nrow(filter(CEDENSURF, Data.source == "CEDEN"))

# Remove duplicate rows of the dataframe using multiple variables

CEDENSURF_DupChecked <- distinct(CEDENSURF, Date, Analyte, CollectionMethod, StationName, Result, .keep_all= TRUE)

nrow(CEDENSURF)-nrow(CEDENSURF_DupChecked)
```
Causes of these records being retained include:

A. No match actually exists in CEDEN. Record SHOULD be retained.

B. CEDEN and SURF have different naming protocols - both Station Name and Station Code differ for the same sites.

C. Latitude and Longitude are rounded differently between the databases.

#### Examples of A

1. There is a measure of endosulfan sulfate at Grizzly Bay in the SURF dataset, but a list of ALL analytes measured at Grizzley Bay in the CEDEN set reveals none including "endosulfan") 

```{r}

# Subset to preview CEDEN Data in the SURF set

head(filter(CEDENSURF, Data.source == "CEDEN"))

# Locate similar location records in CEDEN and SURF using queries

# Record Analyte = "endosulfan sulfate", Agency = Applied Marine Sciences, Inc. California, Collection Method = Filtered water sample, Date = 1997-04-22, Station Name: "Grizzly Bay at Dolphin nr. Suisun Slough. CEDEN: 207SNB0D7", StationCode: "48_52"

# Locate similar location records in CEDEN and SURF using flexible queries

A <- CEDENSURF %>% filter(grepl('Grizzly', StationName)) %>%
  filter(grepl('Dolphin', StationName)) %>%
  filter(Source == "SURF")

B <- CEDENSURF %>% filter(grepl('Grizzly', StationName)) %>%
  filter(grepl('Dolphin', StationName)) %>%
  filter(Source == "CEDEN")

# Sort analytes alphabetically, then display chunk around where "endosulfan" should be
C<- sort(unique(B$Analyte))
C[15:30]
```

2. 
SURF Name: "Toe Drain Nr Babel Slough Nr Freeport Ca"
SURF Code: "57_58"

Could not ID this location in CEDEN data, using grepl() to allow approximate naming. None with "Babel" in StationName


#### Examples of B

It appears that SURF has appended station names from CEDEN with extra info. For example:

SURF: "Grizzly Bay at Dolphin nr. Suisun Slough. CEDEN: 207SNB0D7"
SURF Code: "48_52"

CEDEN: "Grizzly Bay at Dolphin nr. Suisun Slough"
CEDEN Code: "207SNB0D7"

and 

SURF Name: "Sacramento River at Freeport (USGS-11447650)"
SURF Code: "34_5"

CEDEN Name: "Sacramento River at Freeport, CA"
CEDEN Code: 11447650

```{r}
rbind(A[1,], B[1,])
```

#### Example of C

Same station at Grizzly Bay; differences in coordinates listed in SURF (1998 and 2012), compared to CEDEN (2010)

```{r}
C <- CEDENSURF %>%
  filter(grepl('Grizzly', StationName)) %>%
  filter(grepl('Dolphin', StationName)) %>%
  filter(grepl('2012', Date)) %>%
  filter(Data.source == "CEDEN")

rbind(A[1,c(23, 25, 7:10,12)], C[1,c(23, 25, 7:10,12)], B[1,c(23, 25, 7:10,12)])
```

# Next Steps

### 1. Decide whether to retain composite samples, or else how to remove

### 2. Correcting coordinates

Skyler is correcting differences in projection and coordinate data in both CEDEN and SURF

### 3. Naming conventions (analyte, station name, station code, etc)


### Wayne Q: Differences in duplication by year?{.tabset}

There is definitely a more streamlined way to do this, but for expediency... These tables contain the total count of **Exact** duplicates identified in each dataframe. Recall from above that there were more duplicates found usign the final method, but this will give us an initial picture. 

Because there were no exact duplicates for the SURF df, I have not compared their duplication by year.

#### CEDEN WQ
```{r}
CEDENMod_WQ$year <- format(as.Date(CEDENMod_WQ$Date, format="%d/%m/%Y"),"%Y")

aggregate(DupCheck~year,FUN=length,data=CEDENMod_WQ[CEDENMod_WQ$DupCheck=="TRUE",])
```

#### CEDEN Tox

2015 and 2016 were far worse than other years. 

```{r}
CEDENMod_Tox$year <- format(as.Date(CEDENMod_Tox$Date, format="%d/%m/%Y"),"%Y")

aggregate(DupCheck~year,FUN=length,data=CEDENMod_Tox[CEDENMod_Tox$DupCheck=="TRUE",])
```
