---
title: "CEDENSURF_Merge"
author: "Erika W"
date: "2/18/2021"
output:
  html_document:
    code_download: true
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, warning=F, message=F}
library(data.table)
library(lubridate)
library(sf)
library(tidyverse)
```

## Load Data

### CEDEN

Two files - one with tox data, and one with wq data

```{r}
# Load CEDEN Data
CEDENMod_Tox <- fread("https://github.com/WWU-IETC-R-Collab/CEDEN-mod/raw/main/Data/Output/CEDENMod_Toxicity.csv")

CEDENMod_WQ <- fread("https://github.com/WWU-IETC-R-Collab/CEDEN-mod/raw/main/Data/Output/CEDENMod_WQ.csv")
```

### SURF

Two files - one with wq data, and one with SED data ?
```{r}
SURFMod_SED <- fread("https://github.com/WWU-IETC-R-Collab/CEDENSURF-mod/raw/main/Data/Output/SURFMod_SED.csv")

SURFMod_WQ <- fread("https://github.com/WWU-IETC-R-Collab/CEDENSURF-mod/raw/main/Data/Output/SURFMod_water.csv")
```
### Append with source

```{r}
CEDENMod_Tox$Source <- rep("CEDEN", times=nrow(CEDENMod_Tox))

CEDENMod_WQ$Source <- rep("CEDEN", times=nrow(CEDENMod_WQ))

SURFMod_SED$Source <- rep("SURF", times=nrow(SURFMod_SED))

SURFMod_WQ$Source <- rep("SURF", times=nrow(SURFMod_WQ))
```

## Remove duplicate data

More investigating will probably be better, but for now we can simply remove those data defined in SURF as sourced from CEDEN

```{r}

SURFMod_WQ <- filter(SURFMod_WQ, Data.source != "CEDEN")
```

## Merge Water Quality Sets

### Correct mismatched column names

Rather than using Steven's method of renaming using a list (which is dependent on an expected number and order of column names), I renamed each column individually

```{r}
SURFMod_WQ$Unit <- "ppb"

SURFMod_WQ <- SURFMod_WQ %>% rename(Date = Sample_date,
          Analyte = Chemical_name, 
          Result = Concentration..ppb., 
          CollectionMethod = Sample_type, 
          StationCode = Site_code,
          StationName = Site_name,
          MDL = Method_detection_level..ppb.,
          LOQ = Level_of_quantification..ppb.)
```

### Match columns

To do the merge even though there are columns without matches in each dataset, we can give both datasets identical columns and allow them to be filled with NA's when those columns are not actually present.

```{r}
C <- names(CEDENMod_WQ)
S <- names(SURFMod_WQ)

```

Columns that were in SURF and not CEDEN:
`r, diff`

```{r}
#Add missing columns to CEDEN

diff<- setdiff(S, C) # gives items in S that are not in C

CEDENMod_WQ[, diff] <- NA
```

Columns that were in CEDEN and not SURF:
` r, diff`

```{r}
#Add missing columns to SURF

diff<- setdiff(C, S) # gives items in C that are not in S

SURFMod_WQ[, diff] <- NA
```



### Merge

Reorder columns so columns align, then merge with rbind()

```{r}
# Finishing touches before merge; order columns to match - is this really necessary for merge? IDK

C <- sort(names(CEDENMod_WQ))
S <- sort(names(SURFMod_WQ))

SURFMod_WQ <- SURFMod_WQ %>% select(S)

CEDENMod_WQ <- CEDENMod_WQ %>% select(C)

# Check once all columns have perfect matches?
tibble(SURF = names(SURFMod_WQ), CEDEN = names(CEDENMod_WQ))

# MERGE

CEDENSURF_WQ <- rbind(CEDENMod_WQ, SURFMod_WQ)
```

## Merge Joined water data with SURF sediment data


### Correct mismatched column names

Rather than using Steven's method of renaming using a list (which is dependent on an expected number and order of column names), I renamed each column individually

```{r}
SURFMod_SED$Unit <- "ppb"

SURFMod_SED <- SURFMod_SED %>% rename(Date = Sample_date,
          Analyte = Chemical_name, 
          Result = Concentration..ppb., 
          CollectionMethod = Sample_type, 
          StationCode = Site_code,
          StationName = Site_name,
          MDL = Method_detection_level..ppb.,
          LOQ = Level_of_quantification..ppb.)
```

### Match columns

To do the merge even though there are columns without matches in each dataset, we can give both datasets identical columns and allow them to be filled with NA's when those columns are not actually present.

Columns that were in SURF sediment and not in the water tables:
`r, diff`

```{r}

C <- names(CEDENSURF_WQ)
S <- names(SURFMod_SED)

#Add missing columns to CEDEN

diff<- setdiff(S, C) # gives items in S that are not in C

CEDENSURF_WQ[, diff] <- NA
```

Columns that were in the water tables and not in SURF sediment

` r, diff`

```{r}
#Add missing columns to SURF

diff<- setdiff(C, S) # gives items in C that are not in S

SURFMod_SED[, diff] <- NA
```


### Merge

Reorder columns so columns align, then merge with rbind()

```{r}
# Finishing touches before merge; order columns to match - is this really necessary for merge? IDK

C <- sort(names(CEDENSURF_WQ))
S <- sort(names(SURFMod_SED))

SURFMod_SED <- SURFMod_SED %>% select(S)

CEDENSURF_WQ <- CEDENSURF_WQ %>% select(C)

# Check once all columns have perfect matches?

tibble(SURF = names(SURFMod_SED), CEDEN = names(CEDENSURF_WQ))

# MERGE

CEDENSURF_WQSED <- rbind(CEDENSURF_WQ, SURFMod_SED)
```

## Remove duplicate data

An alternate method might be to look for data that match in three specific columns, following the merge; like date, stationname, and result

```{r}
# Ideas from: https://www.datasciencemadesimple.com/remove-duplicate-rows-r-using-dplyr-distinct-function/

# Records in merged data frame
nrow(CEDENSURF_WQSED)

# Remove duplicate rows of the dataframe using multiple variables

CEDENSURF_DupChecked <- distinct(CEDENSURF_WQSED, Date, Analyte, StationName, Result, .keep_all= TRUE)

# Records in checked data
nrow(CEDENSURF_DupChecked)
```
Assuming records that have the exact same station name, date, analyte, and result are duplicates, there were `r nrow(CEDENSURF_WQSED) - nrow(CEDENSURF_DupChecked)` duplicates in the merged data