---
title: "CEDENSURF Merge - Official Short"
author: "Erika W"
date: "2/23/2021"
output:
  html_document:
    code_download: true
    keep_md: true
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=F, message=F}
rm(list = ls())
library(data.table)
library(lubridate)
library(sf)
library(tidyverse)
```

# Intro

This markdown covers the process to identify duplication within the CEDEN and SURF datasets, and to merge those datasets together to make a quality-assured database for use in our analyses.

# Load Data

SURF Data was acquired at the DPR SURF database web page as CSVs via FTP download on 2/17/2021, and modified via the methods outlined in https://github.com/WWU-IETC-R-Collab/CEDENSURF-mod/blob/main/CEDENSURF.md prior to this work.

CEDEN Data was acquired from https://ceden.waterboards.ca.gov/AdvancedQueryTool on January 29 2020 for the Central Valley and San Francisco Bay regions, and spatially queried to the USFE project area. This original data set can be found within the IETC Tox Box at: Upper San Francisco Project\Data & Analyses\Original\CEDEN. The methods of prior modification are at: https://github.com/WWU-IETC-R-Collab/CEDEN-mod/blob/main/CEDEN_ModMaster.md

### CEDEN

Two files - one with tox data, and one with wq data

```{r}
# Load CEDEN Data
CEDENMod_Tox <- fread("https://github.com/WWU-IETC-R-Collab/CEDEN-mod/raw/main/Data/Output/CEDENMod_Toxicity.csv")

CEDENMod_WQ <- fread("https://github.com/WWU-IETC-R-Collab/CEDEN-mod/raw/main/Data/Output/CEDENMod_WQ.csv")
```
```{r}
cSeq<- zoo(CEDENMod_WQ, as.Date(CEDENMod_WQ$Date))

start(cSeq)
end(cSeq)
```
Date range of CEDEN tox data
```{r}
# ID start and end of data range
zSeq <- zoo(CEDENMod_Tox, 
            as.Date(CEDENMod_Tox$Date)) 

start(zSeq)
end(zSeq)
```

### SURF

Two files - one with wq data, and one with sediment data
```{r}
SURFMod_SED <- fread("https://github.com/WWU-IETC-R-Collab/CEDENSURF-mod/raw/main/Data/Output/SURFMod_SED.csv")

SURFMod_WQ <- fread("https://github.com/WWU-IETC-R-Collab/CEDENSURF-mod/raw/main/Data/Output/SURFMod_water.csv")
```
Date range of SURF water data
```{r}
# ID start and end of data range
zSeq <- zoo(SURFMod_WQ, 
            as.Date(SURFMod_WQ$Sample_date)) 

start(zSeq)
end(zSeq)
```
Date range of SURF sediment data
```{r}
# ID start and end of data range
zSeq <- zoo(SURFMod_SED, 
            as.Date(SURFMod_SED$Sample_date)) 

start(zSeq)
end(zSeq)
```

### Append with source

```{r}
CEDENMod_Tox$Source <- rep("CEDEN", times=nrow(CEDENMod_Tox))

CEDENMod_WQ$Source <- rep("CEDEN", times=nrow(CEDENMod_WQ))

SURFMod_SED$Source <- rep("SURF", times=nrow(SURFMod_SED))

SURFMod_WQ$Source <- rep("SURF", times=nrow(SURFMod_WQ))
```

# Remove duplication within each

Due to reported combined efforts to translate CEDEN data to SURF and vice versa, and issues with replicates being retained in each dataset, careful detection and elimination of duplicates should precede any analysis.

## CEDEN Data

There are `r nrow(CEDENMod_WQ)` records in the original WQ dataset

and `r nrow(CEDENMod_Tox)` in the original Tox dataset. 

### CEDEN Tox 

Removing exact duplicates via duplicated() misses duplication of wq data due to multiple species assessments, different sources of data upload, etc. 

Instead, we used distinct() which allows us to assume that records in the same location on the same date,  measuring the same analyte via the same collection method, and obtaining the same result are duplicates. This method deletes almost 50% more records than duplicated() 

```{r}
# Remove duplicate entries, determined as those matching all 5 columns names within distinct().

# Since we are using the Tox database for the water parameters, not the associated organism survival, we remove duplicate WQ entries regardless of the organism assessed.

NoDup_Tox<- distinct(CEDENMod_Tox, Date, StationName, Analyte, CollectionMethod, Result, .keep_all= TRUE) 

# How many duplicate entries were identified and removed?

nrow(CEDENMod_Tox) - nrow(NoDup_Tox)

```

Because of the way we are going to use this data, we are essentially scavenging the WQ parameters associated with each of these tox tests. It is worth considering whether it is worth using this database at all for this purpose;

+ There are only 23 analytes listed, at least 4 are specifically regarding the survival / condition of the organisms being tested

+ Theoretically, the associated WQ data should have already been recorded in the appropriate database.

```{r}
# How many analytes are in this database?
length(unique(NoDup_Tox$Analyte))

# We can also remove records that assess organism status (since we aren't using this for the biotic parameters in our model).

NoDup_Tox <- NoDup_Tox %>% filter(Analyte != "Survival") %>%
  filter(Analyte != "Biomass (wt/orig indiv)") %>%
  filter(Analyte != "Young/female") %>%
  filter(Analyte != "Total Cell Count") %>%
  select(-OrganismName)
```

We're left with only `r nrow(NoDup_Tox)` unique, useful records in the tox dataset - or `r 100*nrow(NoDup_Tox)/nrow(CEDENMod_Tox)` % of the original tox data.

<br>

### CEDEN WQ

Utilizing the distinct() function to assume that records in the same location on the same date, measuring the same analyte via the same collection method and obtaining the same result are duplicates, we find 1536 duplicate records.

That is more than double the number of exact duplicates found, yet still only `r 100*1536/nrow(CEDENMod_WQ)`% of the entire WQ dataset.

```{r}
# Remove duplicate rows of the dataframe using multiple variables

NoDup_WQ <- distinct(CEDENMod_WQ, Date, Analyte, StationName, CollectionMethod, Result, .keep_all= TRUE)

nrow(CEDENMod_WQ) - nrow(NoDup_WQ)

```
<br>

<br>

## Merging CEDEN data

After dealing with duplication WITHIN the CEDEN tox and wq datasets, there were only 9 duplicate records found following the merged data. (75 if Collection Method is not a requirement for establishing duplication)

```{r}
# Vector of column names to compare
WQ <- names(NoDup_WQ)
TOX <- names(NoDup_Tox)

#Add missing columns to CEDEN WQ
DIF<- setdiff(TOX, WQ) # gives items in T that are not in W
NoDup_WQ[, DIF] <- NA

#Add missing columns to CEDEN TOX
DIF<- setdiff(WQ, TOX) # gives items in W that are not in T
NoDup_Tox[, DIF] <- NA

# Finishing touches before merge; order columns to match - is this really necessary for merge? IDK

WQ <- sort(names(NoDup_WQ))
TOX <- sort(names(NoDup_Tox))

NoDup_Tox <- NoDup_Tox %>% select(all_of(TOX))
NoDup_WQ <- NoDup_WQ %>% select(all_of(WQ))

# Check once all columns have perfect matches?
# tibble(SURF = names(NoDup_Tox), CEDEN = names(NoDup_WQ))

# MERGE
CEDEN_ALL <- rbind(NoDup_WQ,NoDup_Tox)
```

### Remove duplicates 

```{r}
# Remove duplicate rows of the dataframe using multiple variables

CEDEN_ALL_DupChecked <- distinct(CEDEN_ALL, Date, Analyte, CollectionMethod, StationName, Result, .keep_all= TRUE)

nrow(CEDEN_ALL)-nrow(CEDEN_ALL_DupChecked)
```

### Further refine

Further assessment revealed problematic data duplication that was not caught when requiring Collection Methods to be equal. We corrected the majority of these errors by:

1. Removing records where Collection Method = "Not Recorded" (Of 305 samples labeled "not recorded" in the entire CEDEN dataset, 153 were duplicated data with non-zero results)

2. (Add another bullet & code if we take action on Sediment Core vs Grab issues)

```{r}
# While there were 307 entries marked "Not Recorded" in the entire combined dataset, 153 were identified as duplicated data. To err on the safe side, we removed all records for which the Collection Method was "Not Recorded"

CEDEN_ALL_DupChecked <- filter(CEDEN_ALL_DupChecked, CollectionMethod != "Not Recorded")

```

### Fix nomenclature

**Split Analyte Column**

Because of formatting differences between the amount of data recorded under "Analyte" in CEDEN compared to the "Chemical_name" in SURF (which will be renamed Analyte), we opted to split the data in CEDEN's analyte column into two columns: 

Analyte (Chemical Name), and Analyte_Type (ie: total or particulate)

Using separate() to split the column and requiring the separation to be a comma and space ", " seems to have worked well, except for one name which appears to be empty

**Convert to lower-case**

SURF chemicals are all lowercase. We can make all letters in the CEDEN data lowercase using tolower() so that they will be compatible.

```{r}

# Split Analyte column

CEDEN_ALL_DupChecked <- CEDEN_ALL_DupChecked %>%
  separate(Analyte, into = c("Analyte", "Analyte_type"), sep = ", " , extra = "merge")

# Convert to lowercase

CEDEN_ALL_DupChecked$Analyte <- tolower(CEDEN_ALL_DupChecked$Analyte)

# Preview
head(sort(unique(CEDEN_ALL_DupChecked$Analyte))) # 908 unique Analytes total

# Looks like requiring the separation from extra to analyte to contain a comma and a space allowed the full names to be retained. Without that, the separation led to an analyte "1" which should have been 1,2-bis(2,4,6- tribromophenoxy)ethane, etc.
```

This simplification of Analyte name would lead to 15,290 more records being considered "duplication" using our current method, mostly (14,776) containing a zero-result. Because these differences in analytes are not retained in the SURF dataset, it makes sense to condense them (remove duplicates) prior to merging the databases.

Removal of these simplified duplicates also eliminates the utility of retaining the "Analyte Type" column. For example, a reading of Analyte X with Type = Total Type = Suspended have the exact same result. Removing duplicates would keep the first record (Analyte = X, Type = Total) and remove the second (Analyte X, Type = S). In the dataframe, if you are trying to reason backwards to the meaning of that remaining meaning (Analyte X, Type = Total), you're missing the other half of the story (Type = Sus too). So, to avoid improper interpretation of this dataframe, Analyte Type should be removed. 

```{r, eval = F}
Check <- distinct(CEDEN_ALL_DupChecked, Date, Analyte, CollectionMethod, StationName, Result, .keep_all= TRUE)

nrow(CEDEN_ALL_DupChecked) - nrow(Check)

DIF<- setdiff(CEDEN_ALL_DupChecked, Check)
length(DIF$Result[DIF$Result == "0"])
```

```{r}
CEDEN_ALL_DupChecked <- distinct(CEDEN_ALL_DupChecked, Date, Analyte, CollectionMethod, StationName, Result, .keep_all= TRUE) %>%
  select(-Analyte_type)
```

### CEDEN merge result

Using these QA/QC methods, `r nrow(CEDEN_ALL_DupChecked)` unique records are available through the CEDEN datasets. 

<br>

<br>

## SURF data

There are `r nrow(SURFMod_WQ)` records in the WQ dataset
and `r nrow(SURFMod_SED)` in the SED dataset. 

There were no exact duplicates in either the WQ or SED data from SURF. Far fewer duplicates were located using our flexible methods than in the CEDEN dataset.

### Data Prep

We renamed columns with analogous data to match CEDEN column names.

```{r}
### SURF WATER

# Move units from embedded in Result column name to their own column
SURFMod_WQ$Unit <- "ppb"

# Rename columns with analogous data to match CEDEN column names.
SURFMod_WQ <- SURFMod_WQ %>% rename(Date = Sample_date,
          Analyte = Chemical_name, 
          Result = Concentration..ppb., 
          CollectionMethod = Sample_type, 
          StationCode = Site_code,
          StationName = Site_name,
          MDL = Method_detection_level..ppb.,
          LOQ = Level_of_quantification..ppb.)

### SURF SEDIMENT

# Move units from embedded in Result column name to their own column
SURFMod_SED$Unit <- "ppb"

# Rename columns with analogous data to match CEDEN column names.
SURFMod_SED <- SURFMod_SED %>% rename(Date = Sample_date,
          Analyte = Chemical_name, 
          Result = Concentration..ppb., 
          CollectionMethod = Sample_type, 
          StationCode = Site_code,
          StationName = Site_name,
          MDL = Method_detection_level..ppb.,
          LOQ = Level_of_quantification..ppb.)
```

### Water: Remove duplicates

Used distinct() to remove duplicates (records in the same location on the same date, measuring the same analyte via the same collection method, and obtaining the same result)

```{r}
# Remove duplicate rows of the dataframe using multiple variables

NoDup_WQ <- distinct(SURFMod_WQ, Date, Analyte, CollectionMethod, StationName, Result, .keep_all= TRUE)
```

We find `r nrow(SURFMod_WQ) - nrow(NoDup_WQ)`, 5,676 of which come from the SURF data sourced from CEDEN

<br>

<br>

### Sediment: Remove duplicates

Used distinct() to remove records in the same location on the same date, measuring the same analyte via the same collection method.

```{r}
# Remove duplicate rows of the dataframe using multiple variables

NoDup_SED <- distinct(SURFMod_SED, Date, Analyte, CollectionMethod, StationName, Result, .keep_all= TRUE)
```

Find only `nrow(SURFMod_SED) - nrow(NoDup_SED)` duplicates, 3336 of which are data sourced from CEDEN

<br>

<br>

## Merge SURF df

```{r}
# Vector of column names to compare
WQ <- names(NoDup_WQ)
SED <- names(NoDup_SED)

#Add missing columns to CEDEN WQ
DIF<- setdiff(SED, WQ) # gives items in S that are not in W
NoDup_WQ[, DIF] <- NA

#Add missing columns to CEDEN SED
DIF<- setdiff(WQ, SED) # gives items in W that are not in S
NoDup_SED[, DIF] <- NA

# Finishing touches before merge; order columns to match - is this really necessary for merge? IDK

WQ <- sort(names(NoDup_WQ))
SED <- sort(names(NoDup_SED))

NoDup_SED <- NoDup_SED %>% select(all_of(SED))
NoDup_WQ <- NoDup_WQ %>% select(all_of(WQ))

# Check once all columns have perfect matches?
# tibble(SURF = names(NoDup_SED), CEDEN = names(NoDup_WQ))

# MERGE
SURF_ALL <- rbind(NoDup_WQ,NoDup_SED)
```

### Duplication between SURF sets

ZERO duplication found between the SED and WQ datasets, assuming duplicates would have to have the exact same Location, Date, Analyte, Collection Method, and Result.

```{r}
# Remove duplicate rows of the dataframe using multiple variables

SURF_ALL_DupChecked <- distinct(SURF_ALL, Date, Analyte, CollectionMethod, StationName, Result, .keep_all= TRUE)

nrow(SURF_ALL)-nrow(SURF_ALL_DupChecked)
```

### Further Refine

We further investigated instances of duplicated entries retained using these methods which differed only in their collection methods, specifically targeting records with identical results != 0.

All but one were corrected by removing records by Study_cd 305, which was an exact replicate of Study_cd 523 except missing collection methods.

```{r}
SURF_ALL_DupChecked <- filter(SURF_ALL_DupChecked, Study_cd != "305")
```

```{r, eval = F}
# Create DF limited to only one record per date,loc,analyte,result combo

Check <- distinct(SURF_ALL, Date, Analyte, StationName, Result, .keep_all= TRUE)

DIF <- setdiff(SURF_ALL_DupChecked, Check)

# How many duplicates of concern are retained after this correction?

length(DIF$Result[DIF$Result != "0"])
```

### Result of SURF Merge

There are `r nrow(SURF_ALL_DupChecked)` unique records available through SURF.

```{r}
# Create a subset of SURF data that excludes data sourced from CEDEN

SURF_ALL_NC <- filter(SURF_ALL_DupChecked, Data.source != "CEDEN")
```

That said, `r nrow(SURF_ALL_DupChecked[SURF_ALL_DupChecked$Data.source == "CEDEN"])` of these records are listed as having been sourced from CEDEN.

In theory only `r nrow(SURF_ALL_NC)` unique records will be contributed through the SURF dataset. Rather than filter these out ahead of the merge, I am retaining them and then using the identification of those records as a test to see whether there are other differentiating factors (such as persisting differences in naming) between the merged dataset that will inhibit our analyses

<br>

<br>

# Merge SURF and CEDEN

Now that each dataset has been independently inspected for duplication *within* each dataset, they can be merged and searched for duplication *between* the datasets.

```{r}
## Data Prep - Match Columns between DF prior to merge

C <- names(CEDEN_ALL_DupChecked)
S <- names(SURF_ALL_DupChecked)

DIF<- setdiff(S, C) # gives items in S that are not in C

#Add missing columns to CEDEN
CEDEN_ALL_DupChecked[, DIF] <- NA

#Add missing columns to SURF
DIF<- setdiff(C, S) # gives items in C that are not in S
SURF_ALL_DupChecked[, DIF] <- NA

# Re-order columns to align
C <- sort(names(CEDEN_ALL_DupChecked)) # 908
S <- sort(names(SURF_ALL_DupChecked)) # 327

SURF_ALL_DupChecked <- SURF_ALL_DupChecked %>% select(all_of(S))
CEDEN_ALL_DupChecked <- CEDEN_ALL_DupChecked %>% select(all_of(C))

# Check once all columns have perfect matches?
# tibble(SURF = names(SURF_ALL_DupChecked), CEDEN = names(CEDEN_ALL_DupChecked))

## MERGE ##

CEDENSURF <- rbind(SURF_ALL_DupChecked, CEDEN_ALL_DupChecked)
```

