---
title: "CEDENSURF_Merge"
author: "Erika W"
date: "2/18/2021"
output:
  html_document:
    code_download: true
    keep_md: true
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=F, message=F}
rm(list = ls())
library(data.table)
library(lubridate)
library(sf)
library(tidyverse)
```

# Load & Prep Data

### CEDEN

Two files - one with tox data, and one with wq data

```{r}
# Load CEDEN Data
CEDENMod_Tox <- fread("https://github.com/WWU-IETC-R-Collab/CEDEN-mod/raw/main/Data/Output/CEDENMod_Toxicity.csv")

CEDENMod_WQ <- fread("https://github.com/WWU-IETC-R-Collab/CEDEN-mod/raw/main/Data/Output/CEDENMod_WQ.csv")
```

### SURF

Two files - one with wq data, and one with SED data ?
```{r}
SURFMod_SED <- fread("https://github.com/WWU-IETC-R-Collab/CEDENSURF-mod/raw/main/Data/Output/SURFMod_SED.csv")

SURFMod_WQ <- fread("https://github.com/WWU-IETC-R-Collab/CEDENSURF-mod/raw/main/Data/Output/SURFMod_water.csv")
```

### Append with source

```{r}
CEDENMod_Tox$Source <- rep("CEDEN", times=nrow(CEDENMod_Tox))

CEDENMod_WQ$Source <- rep("CEDEN", times=nrow(CEDENMod_WQ))

SURFMod_SED$Source <- rep("SURF", times=nrow(SURFMod_SED))

SURFMod_WQ$Source <- rep("SURF", times=nrow(SURFMod_WQ))
```

# Assess duplication within each database

Due to reported combined efforts to translate CEDEN data to SURF and vice versa, and issues with replicates being retained in each dataset, careful detection and elimination of duplicates should precede any analysis.

## CEDEN Data {.tabset}

Due to the structure of the tox data, I have a feeling it involves biological assays that were then related to WQ data sampled on that date (and already present in the CEDEN_WQ dataset. 

There are `r nrow(CEDENMod_WQ)` records in the WQ dataset
and `r nrow(CEDENMod_Tox)` in the Tox dataset. 

### CEDEN Tox 

In the tox dataset, there were several records related to a single species assessment - by creating a group identifier by the location, date, and organism assessed, I calculated that there were 2430 unique samples.

Within the Tox dataset, `r length(unique(CEDENMod_Tox$OrganismName))` different species are represented.

#### Exact duplicates

duplicated() produces a vector identifying which records are duplicates. This can be appended to the original dataset for further investigation. The first record is marked FALSE, while subsequent matches are marked TRUE

```{r}
# add index column identifying which entries are duplicated

CEDENMod_Tox$DupCheck <- duplicated(CEDENMod_Tox)

summary(CEDENMod_Tox$DupCheck)
```

Is it only some programs that have duplication? 9/10 programs do.

```{r}
length(unique(CEDENMod_Tox$Program)) # 10

unique(CEDENMod_Tox$Program[CEDENMod_Tox$DupCheck == "TRUE"])
```

#### Alternative (looser) assessment of duplication

If we assume that some columns may differ due to differences in data loading (ie: originally submitted to SURF and brought to CEDEN vs originally loaded to CEDEN), then we may want to use a looser structure to detect duplicates. 

```{r}
NoDup_Tox<- distinct(CEDENMod_Tox, Date, StationName, Analyte, CollectionMethod, Result, .keep_all= TRUE)

nrow(CEDENMod_Tox) - nrow(NoDup_Tox)
```

By assuming that records in the same location on the same date,  measuring the same analyte via the same collection method, and obtaining the same result are duplicates, we find almost 50% more duplicates. 
25729 records, to be specific. 

That is `r 100*25729/nrow(CEDENMod_Tox)`% of the dataset.

Including Organism name in the list of differentiating variables, it returns almost the exact same count of duplicates (18274) as exact duplication (18297) - indicating that a major source of WQ duplication is repeating WQ parameters for multiple species assessed.

```{r}
# Remove duplicate rows of the dataframe using multiple variables & ORGANISM NAME

nrow(CEDENMod_Tox) - nrow(distinct(CEDENMod_Tox, Date, StationName, CollectionMethod, OrganismName, Analyte, Result, .keep_all= TRUE))
```

Since we are using the df for the water parameters and not the associated organism survival, it's more useful to remove duplicate WQ entries regardless of the organism. We can also remove records that assess 'survival' and 'biomass' (since we aren't using this for the biotic parameters in our model).

```{r}
NoDup_Tox <- NoDup_Tox[NoDup_Tox$Analyte != "Survival"]

NoDup_Tox <- NoDup_Tox[NoDup_Tox$Analyte != "Biomass (wt/orig indiv)"]

# nrow = 31859
```

We're left with only `r nrow(NoDup_Tox)` unique, useful records in the tox dataset - or `r 100*nrow(NoDup_Tox)/nrow(CEDENMod_Tox)` % of the original tox data remaining

<br>

### CEDEN WQ

#### Exact duplicates

duplicated() produces a vector identifying which records are duplicated. This can be appended to the original dataset for further investigation.

```{r}
# add index column identifying which entries are duplicated

CEDENMod_WQ$DupCheck <- duplicated(CEDENMod_WQ)

summary(CEDENMod_WQ$DupCheck)
```

This method identified only 657 exact duplicates in the entire WQ dataset.

Only 9/17 programs present these exact duplicates

```{r}
length(unique(CEDENMod_WQ$Program)) # 17
unique(CEDENMod_WQ$Program[CEDENMod_WQ$DupCheck == "TRUE"])
```

#### Looser assessment of Duplication

Utilizing the distinct() function to assume that records in the same location on the same date, measuring the same analyte via the same collection method and obtaining the same result are duplicates, we find 1536 duplicate records.

That is more than double the number of exact duplicates found, yet still only `r 100*1536/nrow(CEDENMod_WQ)`% of the entire WQ dataset.

```{r}
# Remove duplicate rows of the dataframe using multiple variables

NoDupWQ <- distinct(CEDENMod_WQ, Date, Analyte, StationName, CollectionMethod, Result, .keep_all= TRUE)

nrow(CEDENMod_WQ) - nrow(NoDupWQ)
```
<br>

<br>

## Merging CEDEN data

After dealing with duplication WITHIN the CEDEN tox and wq datasets, there were only 9 duplicate records found following the merged data. (75 if Collection Method is not a requirement for establishing duplication)

```{r}
# Strip column with duplicate check from both, and organism from tox
NoDup_Tox <- select(NoDup_Tox, -c(DupCheck, OrganismName))
NoDupWQ <- select(NoDupWQ, -DupCheck)

# Vector of column names to compare
WQ <- names(NoDupWQ)
TOX <- names(NoDup_Tox)

#Add missing columns to CEDEN WQ
DIF<- setdiff(TOX, WQ) # gives items in T that are not in W
NoDupWQ[, DIF] <- NA

#Add missing columns to CEDEN TOX
DIF<- setdiff(WQ, TOX) # gives items in W that are not in T
NoDup_Tox[, DIF] <- NA

# Finishing touches before merge; order columns to match - is this really necessary for merge? IDK

WQ <- sort(names(NoDupWQ))
TOX <- sort(names(NoDup_Tox))

NoDup_Tox <- NoDup_Tox %>% select(TOX)
NoDupWQ <- NoDupWQ %>% select(WQ)

# Check once all columns have perfect matches?
tibble(SURF = names(NoDup_Tox), CEDEN = names(NoDupWQ))

# MERGE
CEDEN_ALL <- rbind(NoDupWQ,NoDup_Tox)
```

```{r}
# Ideas from: https://www.datasciencemadesimple.com/remove-duplicate-rows-r-using-dplyr-distinct-function/

# Remove duplicate rows of the dataframe using multiple variables

CEDEN_ALL_DupChecked <- distinct(CEDEN_ALL, Date, Analyte, CollectionMethod, StationName, Result, .keep_all= TRUE)

nrow(CEDEN_ALL)-nrow(CEDEN_ALL_DupChecked)
```
### Duplication Errors? {.tabset}

#### Queries
Are there duplicates retained by restricting to unique Collection Methods, which should not be?

There are 401 extra records retained when using Collection method to differentiate identical results. 343 of these duplicate records were not zero-result values, and therefore represent records of concern.

```{r}
CEDEN_ALL_DupChecked <- distinct(CEDEN_ALL, Date, Analyte, 
                                 CollectionMethod, StationName, 
                                 Result, .keep_all= TRUE)

CEDEN_ALL_Check1 <- distinct(CEDEN_ALL, Date, Analyte, 
                              StationName, Result, 
                             .keep_all= TRUE)

nrow(CEDEN_ALL_DupChecked) - nrow(CEDEN_ALL_Check1)
```
```{r}
# ID columns retained only when use Collection Method also

DIF<- setdiff(CEDEN_ALL_DupChecked, CEDEN_ALL_Check1) # gives items in A that are not in B
nrow(DIF)
```
Restricting the duplicates of concern to those which are non-zero values... It appears that there are 6 different collection methods represented in the duplicated data.

Sleuthing what is going on:

1. Many are duplication within a parent project, with one entry having sample method marked "not recorded" and the other defined. 

2. Two projects hold the majority of the duplication: CA Dept of Transit NPDES Monitoring 2014-2015 (153), and San Joaquin County Delta Water Quality Coalition (105)

3. Duplication within the Regional Monitoring Program - Status and Trends deserves closer investigation. Those samples have identical, >0 results for samples labeled with:

+ Collection_Method = Sediment_Core & Matrix = sediment
+ Collection_Method = Sediment_Grab & Matrix = overlyingwater

**--> Question** Does it make sense to have a sample defined as a sediment grab method + matrix of overlying water? Does it make sense for these samples to have identical results?

```{r}
# 343 duplicates had non-zero results

DIF <- DIF %>% filter(., Result != "0") 
nrow(DIF)

# Summary of collection methods which differentiate these duplicates. 

DIF$CollectionMethod <- as.factor(DIF$CollectionMethod)
summary(DIF$CollectionMethod)

# Summary of parent projects represented in the duplicated data

DIF$ParentProject <- as.factor(DIF$ParentProject)
summary(DIF$ParentProject)

# How does number of "not recorded" methods in these duplications compare to those in the entire df?

paste("Collection Method = Not Recorded in...")
tibble(Entire = nrow(CEDEN_ALL_DupChecked[CEDEN_ALL_DupChecked$CollectionMethod== "Not Recorded"]), Dup = nrow(DIF[DIF$CollectionMethod== "Not Recorded"]))


```

#### Printout of duplicated data

Again, these duplicates are retained in the current methodology due to differences between Collection Method. The question is whether they are truly different records, or whether we need to further refine the selection criteria.

```{r}
# Bring back original data associated with each duplicate identified

Dups <- vector("list") # save empty list where each subset be saved

for (i in 1:343){
  Dups[[i]] <- CEDEN_ALL %>% filter(., Analyte == DIF$Analyte[i],
                                         Date == DIF$Date[i], 
                                         StationName == DIF$StationName[i],
                                    Result == DIF$Result[i])
  }

Example <- do.call(rbind, Dups)
Example
```

### Splitting Analyte Column

Because of formatting differences between the amount of data recorded under "Analyte" in CEDEN compared to the "Chemical_name" in SURF (which will be renamed Analyte), we opted to split the data in CEDEN's analyte column into two columns: 
Analyte (Chemical Name), and Analyte_Type (ie: total or particulate)

```{r}
CEDEN_ALL_DupChecked <- CEDEN_ALL_DupChecked %>%
  separate(Analyte, into = c("Analyte", "Analyte_type"), sep = ", " , extra = "merge")

# sort(unique(CEDEN_ALL_DupChecked$Analyte))
head(sort(unique(CEDEN_ALL_DupChecked$Analyte))) # 908 unique Analytes total

# Looks like requiring the separation from extra to analyte to contain a comma and a space allowed the full names to be retained. Without that, the separation led to an analyte "1" which should have been 1,2-bis(2,4,6- tribromophenoxy)ethane, etc.
```

```{r}
# How many records available via full CEDEN merge?
nrow(CEDEN_ALL_DupChecked)
```
In the end, there are 147610 unique records available through the CEDEN datasets. 

<br>

<br>


## SURF data {.tabset}

There are `r nrow(SURFMod_WQ)` records in the WQ dataset
and `r nrow(SURFMod_SED)` in the SED dataset. 

### SURF Water

#### Data Prep

**Correct mismatched column names**

Rather than using Steven's method of renaming using a list (which is dependent on an expected number and order of column names), I renamed each column individually

```{r}
SURFMod_WQ$Unit <- "ppb"

SURFMod_WQ <- SURFMod_WQ %>% rename(Date = Sample_date,
          Analyte = Chemical_name, 
          Result = Concentration..ppb., 
          CollectionMethod = Sample_type, 
          StationCode = Site_code,
          StationName = Site_name,
          MDL = Method_detection_level..ppb.,
          LOQ = Level_of_quantification..ppb.)
```

#### Exact duplicates

duplicated() produces a vector identifying which records are duplicated. This can be appended to the original dataset for further investigation.

```{r}
# add index column identifying which entries are duplicated

SURFMod_WQ$DupCheck <- duplicated(SURFMod_WQ)

summary(SURFMod_WQ$DupCheck)
```

Fantastic. There are no exact duplicates within the SURF water data - even when we include the CEDEN allowed in SURF

#### Looser assessment of Duplication

Utilizing the distinct() function to assume that records in the same location on the same date, measuring the same analyte and obtaining the same result are duplicates, we find 17756 (there were only 5,676 w/o CEDEN data) duplicate records.

```{r}
# Remove duplicate rows of the dataframe using multiple variables

NoDupWQ <- distinct(SURFMod_WQ, Date, Analyte, CollectionMethod, StationName, Result, .keep_all= TRUE)

nrow(SURFMod_WQ) - nrow(NoDupWQ)
```
<br>

<br>

### SURF sediment

#### Data Prep

More investigating will probably be better, but for now we can simply remove those data defined in SURF as sourced from CEDEN

**Correct mismatched column names**
```{r}
SURFMod_SED$Unit <- "ppb"

SURFMod_SED <- SURFMod_SED %>% rename(Date = Sample_date,
          Analyte = Chemical_name, 
          Result = Concentration..ppb., 
          CollectionMethod = Sample_type, 
          StationCode = Site_code,
          StationName = Site_name,
          MDL = Method_detection_level..ppb.,
          LOQ = Level_of_quantification..ppb.)
```

#### Exact duplicates

duplicated() produces a vector identifying which records are duplicated. This can be appended to the original dataset for further investigation.

```{r}
# add index column identifying which entries are duplicated

SURFMod_SED$DupCheck <- duplicated(SURFMod_SED)

summary(SURFMod_SED$DupCheck)
```

Fantastic. There are no exact duplicates within the SURF sediment data, even with CEDEN also contained.

#### Looser assessment of Duplication

Utilizing the distinct() function to assume that records in the same location on the same date, measuring the same analyte and obtaining the same result are duplicates, we find only 4,120 (was 784 without CEDEN sources)

```{r}
# Remove duplicate rows of the dataframe using multiple variables

NoDupSED <- distinct(SURFMod_SED, Date, Analyte, CollectionMethod, StationName, Result, .keep_all= TRUE)

nrow(SURFMod_SED) - nrow(NoDupSED)
```
<br>

<br>

## Merge SURF df

```{r}
# Strip column with duplicate check from both
NoDupSED <- select(NoDupSED, -DupCheck)
NoDupWQ <- select(NoDupWQ, -DupCheck)

# Vector of column names to compare
WQ <- names(NoDupWQ)
SED <- names(NoDupSED)

#Add missing columns to CEDEN WQ
DIF<- setdiff(SED, WQ) # gives items in S that are not in W
NoDupWQ[, DIF] <- NA

#Add missing columns to CEDEN SED
DIF<- setdiff(WQ, SED) # gives items in W that are not in S
NoDupSED[, DIF] <- NA

# Finishing touches before merge; order columns to match - is this really necessary for merge? IDK

WQ <- sort(names(NoDupWQ))
SED <- sort(names(NoDupSED))

NoDupSED <- NoDupSED %>% select(SED)
NoDupWQ <- NoDupWQ %>% select(WQ)

# Check once all columns have perfect matches?
tibble(SURF = names(NoDupSED), CEDEN = names(NoDupWQ))

# MERGE
SURF_ALL <- rbind(NoDupWQ,NoDupSED)
```

### Duplication between SURF sets

ZERO duplication found between the SED and WQ datasets, assuming duplicates would have to have the exact same Location, Date, Analyte, Collection Method, and Result.

```{r}
# Ideas from: https://www.datasciencemadesimple.com/remove-duplicate-rows-r-using-dplyr-distinct-function/

# Remove duplicate rows of the dataframe using multiple variables

SURF_ALL_DupChecked <- distinct(SURF_ALL, Date, Analyte, CollectionMethod, StationName, Result, .keep_all= TRUE)

nrow(SURF_ALL)-nrow(SURF_ALL_DupChecked)
```

### Request: Examples of SURF duplicates when Collection Method is omitted? {.tabset}

#### With CEDEN Data

Able to make df of "duplicate" entries for further inspection. Though **ALL WERE DUE TO RESULT == 0 WITHOUT CEDEN INCLUDED**

When retaining data sourced from CEDEN, there are some instances of results != 0 with identical results.

```{r}
# Create DF limited to only one record per date,loc,analyte,result combo

SURF_ALL_Check1 <- distinct(SURF_ALL, Date, Analyte, StationName, Result, .keep_all= TRUE)

# ID columns retained only when use Collection Method also

DIF<- setdiff(SURF_ALL_DupChecked, SURF_ALL_Check1) # gives items in A that are not in B

# Produce vector of all different "Result" values in the duplicated data

length(unique(SURF_ALL$Result)) # how many different "result" values are represented in the entire combined SURF dataset

unique(DIF$Result) # All unique "result" values within the duplicates that were removed.
```

These duplicates have some commonalities;

1. they are all brought into SURF from "CEDEN" (except for one "OTHER")

2. they have different "collection methods", typically one with a collection method specified (Study CD 523), and one without the method specified (Study CD 305)

3. Duplicates are almost all between Study CD's 523 & 305. Only one pair of replicates were due to Study CD's 451 & 80. 

```{r}

DIF <- DIF %>% filter(., Result != "0") # Only 19 duplicates were removed that had non-0-result values

# Bring back original data associated with each duplicate identified

Dups <- vector("list") # save empty list where each subset be saved

for (i in 1:19){
  Dups[[i]] <- SURF_ALL %>% filter(., Analyte == DIF$Analyte[i],
                                         Date == DIF$Date[i], 
                                         StationName == DIF$StationName[i])
  }

Example <- do.call(rbind, Dups)
Example
```

```{r}
unique(Example$Data.source)

unique(Example$Study_cd)
```

Given this info, it is worthwhile to ask if there are any unique entries by Study_cd 305, or if all are duplications of more complete data by Study_523.

**ALL ENTRIES IN SURF BY STUDY CD 305 ARE DUPS**

```{r}

length(SURF_ALL$Study_cd[SURF_ALL$Study_cd == "305"]) # Number of entries by this Agency/Program in all SURF data

length(Example$Study_cd[SURF_ALL$Study_cd == "305"]) # Number of entries that were part of the duplication findings
```

#### Without CEDEN 

ZERO duplication found between the SED and WQ datasets, assuming duplicates would have to have the exact same Location, Date, Analyte, Collection Method, and Result.

```{r}
# Ideas from: https://www.datasciencemadesimple.com/remove-duplicate-rows-r-using-dplyr-distinct-function/

# Filter out data sourced from CEDEN

SURF_ALL_NC <- filter(SURF_ALL, Data.source != "CEDEN")
# Remove duplicate rows of the dataframe using multiple variables

SURF_ALL_DupCheckedNC <- distinct(SURF_ALL_NC, Date, Analyte, CollectionMethod, StationName, Result, .keep_all= TRUE)

nrow(SURF_ALL_NC)-nrow(SURF_ALL_DupCheckedNC)
```

*Note* I had started without 'Collection Method' in the distinct() query, but added it as an additional requirement for establishing duplication because without that record, the difference between results for water vs sediment were not recognized and **26127** records were considered duplicates following the merging of the SURF data.

This is still surprising to me, because that means the RESULT concentration was identical for both sediment and water samples. 

By creating a df of "duplicate" entries for further inspection. Findings confirm that
**ALL ARE DUE TO RESULT == 0**

```{r}
# Create DF limited to only one record per date,loc,analyte,result combo

SURF_ALL_Check1 <- distinct(SURF_ALL_NC, Date, Analyte, StationName, Result, .keep_all= TRUE)

# ID columns retained only when use Collection Method also

DIF<- setdiff(SURF_ALL_DupCheckedNC, SURF_ALL_Check1) # gives items in A that are not in B

# Produce vector of all different "Result" values in the duplicated data

length(unique(SURF_ALL_NC$Result)) # how many different "result" values are represented in the entire combined SURF dataset

unique(DIF$Result) # All unique "result" values within the duplicates that were removed.
```

### Result of SURF Merge

Due to duplication found above, remove study 305 that was entirely duplication prior to next steps (Because this duplication was not able to be removed by simply filtering on date, loc, analyte, collection method, and result)

```{r}
SURF_ALL_DupChecked <- filter(SURF_ALL_DupChecked, Study_cd != "305")

nrow(SURF_ALL_DupChecked)
```
There are 143,000 unique records available through SURF.

<br>

<br>

# Merge SURF and CEDEN
Now that each dataset has been independently inspected for duplication *within* each dataset, they can be merged and searched for duplication *between* the datasets.

### Data prep

**Match columns to CEDEN data**

```{r}
C <- names(CEDEN_ALL_DupChecked)
S <- names(SURF_ALL_DupChecked)

DIF<- setdiff(S, C) # gives items in S that are not in C
```

Columns that were in SURF and not CEDEN:
`r DIF`

```{r}
#Add missing columns to CEDEN
CEDEN_ALL_DupChecked[, DIF] <- NA
```

```{r}
#Add missing columns to SURF
DIF<- setdiff(C, S) # gives items in C that are not in S
SURF_ALL_DupChecked[, DIF] <- NA
```

Columns that were in CEDEN and not SURF:
` r DIF`

### Merge

Reorder columns to align, then merge with rbind()

```{r}
# Finishing touches before merge; order columns to match - is this really necessary for merge? IDK

C <- sort(names(CEDEN_ALL_DupChecked)) # 908
S <- sort(names(SURF_ALL_DupChecked)) # 327

SURF_ALL_DupChecked <- SURF_ALL_DupChecked %>% select(S)

CEDEN_ALL_DupChecked <- CEDEN_ALL_DupChecked %>% select(C)

# Check once all columns have perfect matches?
tibble(SURF = names(SURF_ALL_DupChecked), CEDEN = names(CEDEN_ALL_DupChecked))

# MERGE
CEDENSURF <- rbind(SURF_ALL_DupChecked, CEDEN_ALL_DupChecked)
```

### Check for duplicates

None found! Great start.

```{r}
# Ideas from: https://www.datasciencemadesimple.com/remove-duplicate-rows-r-using-dplyr-distinct-function/

# Remove duplicate rows of the dataframe using multiple variables

CEDENSURF_DupChecked <- distinct(CEDENSURF, Date, Analyte, CollectionMethod, StationName, Result, .keep_all= TRUE)

nrow(CEDENSURF)-nrow(CEDENSURF_DupChecked)
```
# Next Steps

### Naming conventions

We should check whether splitting the Analyte column in CEDEN helped to unify the nomenclature between datasets, before we can confidently suggest there is no more duplication.

The lists of analytes are MUCH closer to eachother using this method, though it isn't perfect. Some further generalizations seem to exist in SURF Chemical_names, and more analytes are recorded in CEDEN
```{r}
analyte_C <- sort(unique(CEDENSURF$Analyte[CEDENSURF$Source == "CEDEN"]))
# 908 (was >1800 before splitting the column)


analyte_S <- sort(unique(CEDENSURF$Analyte[CEDENSURF$Source == "SURF"]))
# 307

DIF <- setdiff(analyte_C, analyte_S)
# Main things in C and not in S:
# PCH as ... three dif, 
# PCB - MANY differnt numbers, 
# PBDE = MANY
# Dechlorane - several
# Two each for DDD, DDE, and DDT

# Funny analytes to note: Young/female
```

There are ` r length(analyte_C) ` different analytes in the CEDEN data, while only `r length(analyte_S)` in the SURF.

It appears that SURF gives the simple analyte name, while CEDEN retains additional information in the analyte name specifying 'total' vs 'particulate' vs 
'dissolved'.

For the purposes of this merged dataset, it may make sense to simplify the analyte names from CEDEN, and allow the original (unmerged with SURF) CEDEN data to be used in situations where that additional information is required.

For example, a subset of "Analytes" in each database follow:
```{r}
# Show example of differences
tibble(CEDEN = c(analyte_C[20:40]), SURF = c(analyte_S[10:30]))
```

### Wayne Q: Differences in duplication by year?{.tabset}

There is definitely a more streamlined way to do this, but for expediency... These tables contain the total count of **Exact** duplicates identified in each dataframe. Recall from above that there were more duplicates found usign the final method, but this will give us an initial picture. 

Because there were no exact duplicates for the SURF df, I have not compared their duplication by year.

#### CEDEN WQ
```{r}
CEDENMod_WQ$year <- format(as.Date(CEDENMod_WQ$Date, format="%d/%m/%Y"),"%Y")

aggregate(DupCheck~year,FUN=length,data=CEDENMod_WQ[CEDENMod_WQ$DupCheck=="TRUE",])
```

#### CEDEN Tox

2015 and 2016 were far worse than other years. 

```{r}
CEDENMod_Tox$year <- format(as.Date(CEDENMod_Tox$Date, format="%d/%m/%Y"),"%Y")

aggregate(DupCheck~year,FUN=length,data=CEDENMod_Tox[CEDENMod_Tox$DupCheck=="TRUE",])
```
