---
title: "CEDENSURF_Merge"
author: "Erika W"
date: "2/18/2021"
output:
  html_document:
    code_download: true
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=F, message=F}
library(data.table)
library(lubridate)
library(sf)
library(tidyverse)
```

## Load Data

### CEDEN

Two files - one with tox data, and one with wq data

```{r}
# Load CEDEN Data
CEDENMod_Tox <- fread("https://github.com/WWU-IETC-R-Collab/CEDEN-mod/raw/main/Data/Output/CEDENMod_Toxicity.csv")

CEDENMod_WQ <- fread("https://github.com/WWU-IETC-R-Collab/CEDEN-mod/raw/main/Data/Output/CEDENMod_WQ.csv")
```

### SURF

Two files - one with wq data, and one with SED data ?
```{r}
SURFMod_SED <- fread("https://github.com/WWU-IETC-R-Collab/CEDENSURF-mod/raw/main/Data/Output/SURFMod_SED.csv")

SURFMod_WQ <- fread("https://github.com/WWU-IETC-R-Collab/CEDENSURF-mod/raw/main/Data/Output/SURFMod_water.csv")
```
### Append with source

```{r}
CEDENMod_Tox$Source <- rep("CEDEN", times=nrow(CEDENMod_Tox))

CEDENMod_WQ$Source <- rep("CEDEN", times=nrow(CEDENMod_WQ))

SURFMod_SED$Source <- rep("SURF", times=nrow(SURFMod_SED))

SURFMod_WQ$Source <- rep("SURF", times=nrow(SURFMod_WQ))
```

## Remove duplicate data

More investigating will probably be better, but for now we can simply remove those data defined in SURF as sourced from CEDEN

```{r}
SURFMod_WQ <- filter(SURFMod_WQ, Data.source != "CEDEN")
```

## Merge Water Quality Sets

### Correct mismatched column names

Rather than using Steven's method of renaming using a list (which is dependent on an expected number and order of column names), I renamed each column individually

```{r}
SURFMod_WQ$Unit <- "ppb"

SURFMod_WQ <- SURFMod_WQ %>% rename(Date = Sample_date,
          Analyte = Chemical_name, 
          Result = Concentration..ppb., 
          CollectionMethod = Sample_type, 
          StationCode = Site_code,
          StationName = Site_name,
          MDL = Method_detection_level..ppb.,
          LOQ = Level_of_quantification..ppb.)
```

### Match columns

To do the merge even though there are columns without matches in each dataset, we can give both datasets identical columns and allow them to be filled with NA's when those columns are not actually present.

```{r}
C <- names(CEDENMod_WQ)
S <- names(SURFMod_WQ)

DIF<- setdiff(S, C) # gives items in S that are not in C
```

Columns that were in SURF and not CEDEN:
`r DIF`

```{r}
#Add missing columns to CEDEN

CEDENMod_WQ[, DIF] <- NA
```


```{r}
#Add missing columns to SURF

DIF<- setdiff(C, S) # gives items in C that are not in S

SURFMod_WQ[, DIF] <- NA
```

Columns that were in CEDEN and not SURF:
` r DIF`

### Merge

Reorder columns so columns align, then merge with rbind()

```{r}
# Finishing touches before merge; order columns to match - is this really necessary for merge? IDK

C <- sort(names(CEDENMod_WQ))
S <- sort(names(SURFMod_WQ))

SURFMod_WQ <- SURFMod_WQ %>% select(S)

CEDENMod_WQ <- CEDENMod_WQ %>% select(C)

# Check once all columns have perfect matches?
tibble(SURF = names(SURFMod_WQ), CEDEN = names(CEDENMod_WQ))

# MERGE

CEDENSURF_WQ <- rbind(CEDENMod_WQ, SURFMod_WQ)
```

## Merge Joined water data with SURF sediment data


### Correct mismatched column names

Rather than using Steven's method of renaming using a list (which is dependent on an expected number and order of column names), I renamed each column individually

```{r}
SURFMod_SED$Unit <- "ppb"

SURFMod_SED <- SURFMod_SED %>% rename(Date = Sample_date,
          Analyte = Chemical_name, 
          Result = Concentration..ppb., 
          CollectionMethod = Sample_type, 
          StationCode = Site_code,
          StationName = Site_name,
          MDL = Method_detection_level..ppb.,
          LOQ = Level_of_quantification..ppb.)
```

### Match columns

To do the merge even though there are columns without matches in each dataset, we can give both datasets identical columns and allow them to be filled with NA's when those columns are not actually present.

```{r}

C <- names(CEDENSURF_WQ)
S <- names(SURFMod_SED)

#Add missing columns to CEDEN

DIF<- setdiff(S, C) # gives items in S that are not in C

CEDENSURF_WQ[, DIF] <- NA
```

Columns that were in SURF sediment and not in the water tables:
`r DIF`


```{r}
#Add missing columns to SURF

DIF<- setdiff(C, S) # gives items in C that are not in S

SURFMod_SED[, DIF] <- NA
```


Columns that were in the water tables and not in SURF sediment:

`r DIF`

### Merge

Reorder columns so columns align, then merge with rbind()

```{r}
# Finishing touches before merge; order columns to match - is this really necessary for merge? IDK

C <- sort(names(CEDENSURF_WQ))
S <- sort(names(SURFMod_SED))

SURFMod_SED <- SURFMod_SED %>% select(S)

CEDENSURF_WQ <- CEDENSURF_WQ %>% select(C)

# Check once all columns have perfect matches?

tibble(SURF = names(SURFMod_SED), CEDEN = names(CEDENSURF_WQ))

# MERGE

CEDENSURF_WQSED <- rbind(CEDENSURF_WQ, SURFMod_SED)
```

## Remove duplicate data

### Removal from final merged document

distinct() allows us to remove data that matches in specific columns, following the merge; like Date, StationName, Analyte, and Result. It keeps the first record and removes all subsequent matches. 

```{r}
# Ideas from: https://www.datasciencemadesimple.com/remove-duplicate-rows-r-using-dplyr-distinct-function/

# Records in merged data frame
nrow(CEDENSURF_WQSED)

# Remove duplicate rows of the dataframe using multiple variables

CEDENSURF_DupChecked <- distinct(CEDENSURF_WQSED, Date, Analyte, StationName, Result, .keep_all= TRUE)

# Records in checked data
nrow(CEDENSURF_DupChecked)
```

Assuming records that have the exact same station name, date, analyte, and result are duplicates, there were `r nrow(CEDENSURF_WQSED) - nrow(CEDENSURF_DupChecked)` duplicates in the merged data

< br >

## Investigating data first (late)

Due to the structure of the tox data, I have a feeling it involves biological assays that were then related to WQ data sampled on that date (and already present in the CEDEN_WQ dataset)

I can look to see if this is true by merging the two ceden datasets following the same method as before, and assessing differences in sample length. 

### Original Data {.tabset}

```{r}
# Load CEDEN Data
CEDENMod_Tox <- fread("https://github.com/WWU-IETC-R-Collab/CEDEN-mod/raw/main/Data/Output/CEDENMod_Toxicity.csv")

CEDENMod_WQ <- fread("https://github.com/WWU-IETC-R-Collab/CEDEN-mod/raw/main/Data/Output/CEDENMod_WQ.csv")
```

There are `r nrow(CEDENMod_WQ)` records in the WQ dataset
and `r nrow(CEDENMod_Tox)` in the Tox dataset. 

#### Tox Data

In the tox dataset, there were several records related to a single species assessment - by creating a group identifier by the location, date, and organism assessed, I calculated that there were 2430 unique samples.

Within the Tox dataset, `r length(unique(CEDENMod_Tox$OrganismName))` different species are represented.

**Duplication within Tox data**

Does the method of assuming differences *between the datasets* detect anomalies within the dataset? YES - 25989 records, to be specific.

```{r}
# Remove duplicate rows of the dataframe using multiple variables

Check <- distinct(CEDENMod_Tox, Date, Analyte, StationName, Result, .keep_all= TRUE)

nrow(CEDENMod_Tox) - nrow(Check)

Check <- distinct(CEDENMod_Tox, Date, StationName, OrganismName, Analyte, Result, .keep_all= TRUE)

nrow(CEDENMod_Tox) - nrow(Check)
```
Even assuming that multiple species may be checked on the same day, so WQ parameters would be replicated for each species assessed on that day, there are still `r nrow(CEDENMod_Tox) - nrow(Check)` duplicates, which should not be included.

#### WQ Data

**Duplication within WQ data**

Does the method of assuming differences *between the datasets* detect anomalies within the dataset? YES - 1719 records, to be specific.

```{r}
# Remove duplicate rows of the dataframe using multiple variables

Check <- distinct(CEDENMod_WQ, Date, Analyte, StationName, Result, .keep_all= TRUE)

nrow(CEDENMod_WQ) - nrow(Check)
```


### Repeat Merge with only CEDEN data frames

Due to the detections described above, further sleuthing should be done before using this merge. But, for the record, here would be the process.

```{r}
WQ <- names(CEDENMod_WQ)
TOX <- names(CEDENMod_Tox)

DIF<- setdiff(TOX, WQ) # gives items in T that are not in W

#Add missing columns to CEDEN WQ

CEDENMod_WQ[, DIF] <- NA

#Add missing columns to CEDEN TOX

DIF<- setdiff(WQ, TOX) # gives items in W that are not in T

CEDENMod_Tox[, DIF] <- NA

# Finishing touches before merge; order columns to match - is this really necessary for merge? IDK

WQ <- sort(names(CEDENMod_WQ))
TOX <- sort(names(CEDENMod_Tox))

CEDENMod_Tox <- CEDENMod_Tox %>% select(TOX)

CEDENMod_WQ <- CEDENMod_WQ %>% select(WQ)

# Check once all columns have perfect matches?
tibble(SURF = names(CEDENMod_Tox), CEDEN = names(CEDENMod_WQ))

# MERGE

CEDEN_ALL <- rbind(CEDENMod_WQ,CEDENMod_Tox)
```

### Detect differences

```{r}
CEDENMod_Tox$ID <- paste(CEDENMod_Tox$Date, CEDENMod_Tox$StationCode, CEDENMod_Tox$OrganismName, sep= ",")

length(unique(CEDENMod_Tox$ID))
```
```{r}
# Ideas from: https://www.datasciencemadesimple.com/remove-duplicate-rows-r-using-dplyr-distinct-function/

# Remove duplicate rows of the dataframe using multiple variables

CEDEN_ALL_DupChecked <- distinct(CEDEN_ALL, Date, Analyte, StationName, Result, .keep_all= TRUE)

nrow(CEDEN_ALL_DupChecked)
```
Assuming records that have the exact same station name, date, analyte, and result are duplicates, there were `r nrow(CEDEN_ALL) - nrow(CEDEN_ALL_DupChecked)` duplicates in the merged data, leaving `r nrow(CEDEN_ALL_DupChecked)` total records in the merged df.

That suggests that only `r nrow(CEDEN_ALL_DupChecked) - nrow(CEDENMod_WQ)` unique records were brought over from the tox dataset

### Further refining - what is left?

If we remove records that assess 'survival' and 'biomass' (since we aren't using this for biotic parameters in our model)...

```{r}
CEDEN_ALL_DupChecked <- CEDEN_ALL_DupChecked[CEDEN_ALL_DupChecked$Analyte != "Survival"]

CEDEN_ALL_DupChecked <- CEDEN_ALL_DupChecked[CEDEN_ALL_DupChecked$Analyte != "Biomass (wt/orig indiv)"]
```

We're left with `r nrow(CEDEN_ALL_DupChecked) - nrow(CEDENMod_WQ)` unique records brought over out of all `r nrow(CEDENMod_Tox)` records in the Tox dataset 
