---
title: "Analyte Names"
author: "Erika W"
date: "2/18/2021"
output:
  html_document:
    code_download: true
    keep_md: true
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=F, message=F}
rm(list = ls())
library(data.table)
library(lubridate)
library(sf)
library(tidyverse)
```

## Load Data

### CEDEN

Two files - one with tox data, and one with wq data

```{r}
# Load CEDEN Data
CEDENMod_Tox <- fread("https://github.com/WWU-IETC-R-Collab/CEDEN-mod/raw/main/Data/Output/CEDENMod_Toxicity.csv")

CEDENMod_WQ <- fread("https://github.com/WWU-IETC-R-Collab/CEDEN-mod/raw/main/Data/Output/CEDENMod_WQ.csv")
```

### SURF

Two files - one with wq data, and one with sediment data
```{r}
SURFMod_SED <- fread("https://github.com/WWU-IETC-R-Collab/CEDENSURF-mod/raw/main/Data/Output/SURFMod_SED.csv")

SURFMod_WQ <- fread("https://github.com/WWU-IETC-R-Collab/CEDENSURF-mod/raw/main/Data/Output/SURFMod_water.csv")
```

### Append with source

```{r}
CEDENMod_Tox$Source <- rep("CEDEN", times=nrow(CEDENMod_Tox))

CEDENMod_WQ$Source <- rep("CEDEN", times=nrow(CEDENMod_WQ))

SURFMod_SED$Source <- rep("SURF", times=nrow(SURFMod_SED))

SURFMod_WQ$Source <- rep("SURF", times=nrow(SURFMod_WQ))
```

## Merge within source {.tabset}

To facilitate ID of analytes in each source, I merged the CEDEN sets here and the SURF sets (WITHOUT ADDRESSING DUPLICATION - since that shouldn't matter for these purposes. 

If you want to see CEDEN tox vs wq analyte lists separately, we can also do that.

### CEDEN
```{r}
# Vector of column names to compare
WQ <- names(CEDENMod_WQ)
TOX <- names(CEDENMod_Tox)

#Add missing columns to CEDEN WQ
DIF<- setdiff(TOX, WQ) # gives items in T that are not in W
CEDENMod_WQ[, DIF] <- NA

#Add missing columns to CEDEN TOX
DIF<- setdiff(WQ, TOX) # gives items in W that are not in T
CEDENMod_Tox[, DIF] <- NA

# Finishing touches before merge; order columns to match - is this really necessary for merge? IDK

WQ <- sort(names(CEDENMod_WQ))
TOX <- sort(names(CEDENMod_Tox))

CEDENMod_Tox <- CEDENMod_Tox %>% select(TOX)
CEDENMod_WQ <- CEDENMod_WQ %>% select(WQ)

# If rbind doesnt work, Check here to ensure columns have perfect matches
# tibble(Tox = names(CEDENMod_Tox), WQ = names(CEDENMod_WQ))

# MERGE
CEDEN_ALL <- rbind(CEDENMod_WQ,CEDENMod_Tox)
```

### SURF
```{r}
# Vector of column names to compare
WQ <- names(SURFMod_WQ)
SED <- names(SURFMod_SED)

#Add missing columns to SURF WQ
DIF<- setdiff(SED, WQ) # gives items in T that are not in W
SURFMod_WQ[, DIF] <- NA

#Add missing columns to SURF SED
DIF<- setdiff(WQ, SED) # gives items in W that are not in T
SURFMod_SED[, DIF] <- NA

# Finishing touches before merge; order columns to match - is this really necessary for merge? IDK

WQ <- sort(names(SURFMod_WQ))
SED <- sort(names(SURFMod_SED))

SURFMod_SED <- SURFMod_SED %>% select(SED)
SURFMod_WQ <- SURFMod_WQ %>% select(WQ)

# If rbind doesnt work, Check here to ensure columns have perfect matches
# tibble(SED = names(SURFMod_SED), WQ = names(SURFMod_WQ))

# MERGE
SURF_ALL <- rbind(SURFMod_WQ,SURFMod_SED)

# Correct SURF column names to match CEDEN
SURF_ALL$Unit <- "ppb"

SURF_ALL <- SURF_ALL %>% rename(Date = Sample_date,
          Analyte = Chemical_name, 
          Result = Concentration..ppb., 
          CollectionMethod = Sample_type, 
          StationCode = Site_code,
          StationName = Site_name,
          MDL = Method_detection_level..ppb.,
          LOQ = Level_of_quantification..ppb.)
```

## Fixing CEDEN nomenclature

**Split Analyte Column**

Because of formatting differences between the amount of data recorded under "Analyte" in CEDEN compared to the "Chemical_name" in SURF (which will be renamed Analyte), we opted to split the data in CEDEN's analyte column into two columns: 

Analyte (Chemical Name), and Analyte_Type (ie: total or particulate)

Using separate() to split the column and requiring the separation to be a comma and space ", " seems to have worked well, except for one name which appears to be empty

**Convert to lower-case**

SURF chemicals are all lowercase. We can make all letters in the CEDEN data lowercase using tolower() so that they will be compatible.

```{r}

# Split Analyte column

CEDEN_ALL_Renamed <- CEDEN_ALL %>%
  separate(Analyte, into = c("Analyte", "Analyte_type"), sep = ", " , extra = "merge")

# Convert to lowercase

CEDEN_ALL_Renamed$Analyte <- tolower(CEDEN_ALL_Renamed$Analyte)

# Preview
head(sort(unique(CEDEN_ALL_Renamed$Analyte))) # 908 unique Analytes total

# Looks like requiring the separation from extra to analyte to contain a comma and a space allowed the full names to be retained. Without that, the separation led to an analyte "1" which should have been 1,2-bis(2,4,6- tribromophenoxy)ethane, etc.
```

## Unique analyte names

### Identify unique analyte names

Create vectors that contain unique analyte names from each DF; sorted alphabetically

```{r}
CEDEN_Orig_Analytes <- sort(unique(CEDEN_ALL$Analyte)) #1810 unique analytes before we split the names from modifying text

CEDEN_Analytes <- sort(unique(CEDEN_ALL_Renamed$Analyte)) #757 unique analytes

SURF_Analytes <- sort(unique(SURF_ALL$Analyte)) #327
```

You can also look at analytes that are specifically in CEDEN and not in SURF using this code:

```{r}
DIF <- setdiff(CEDEN_Analytes, SURF_Analytes)

## Print DIF to see analytes in C and not in S

# Main things I noticed....
# PCH as ... three dif, 
# PCB - MANY differnt numbers, 
# PBDE = MANY
# Dechlorane - several
# Two each for DDD, DDE, and DDT

# Funny analytes to note: Young/female
```

### Create table of names {.tabset}

Both these output tables use the "corrected" naming of the CEDEN data. If you want to compare the original names, the vector "CEDEN_Orig_Analytes" contains those.


#### Format 1

If you want a single column of analyte names, and another identifying source (CEDEN vs SURF) run this code and remove the hashtag to write an output table.

```{r}
# DF of unique analytes for each

CE <- distinct(CEDEN_ALL_Renamed, Analyte, Source)
CE$Analyte <- CEDEN_Analytes # Easiest way to rearrange them alphabetically

SE <- distinct(SURF_ALL, Analyte, Source)
SE$Analyte <- SURF_Analytes # Easiest way to rearrange them alphabetically

#Unite into one DF
Analytes <- rbind(CE,SE)

# Run this code to save this file
write.csv(x = Analytes, file = "CEDENSURF_Analytes_Format1.csv", row.names = F)
```

#### Format 2

If you want a table of SURF and CEDEN analytes in adjacent columns like below, run this code and remove the hashtag to write an output table.

```{r}
# Example of differences
tibble(CEDEN = CEDEN_Analytes[2:11], SURF = SURF_Analytes[1:10])

# Create empty DF with correct dimensions
Analytes2 <- setNames(data.frame(matrix(ncol = 2, nrow = 757)), c("CEDEN", "SURF"))

# Add in data
Analytes2$CEDEN <- CEDEN_Analytes

Analytes2$SURF[1:327] <- SURF_Analytes

# Run this code to save this file
# write.csv(x = Analytes2, file = "CEDENSURF_Analytes_Format2.csv", row.names = F)
```
### Example grepl() name query 

```{r}
Analytes[grepl('nitra', Analyte)]
```

