---
title: "IssuesDocumentation"
author: "Erika W"
date: "3/2/2021"
output:
  html_document:
    code_download: true
    keep_md: true
    code_folding: hide
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=F, message=F}
rm(list=setdiff(ls(), c("CEDENSURF")))

setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

library(data.table)
library(lubridate)
library(sf)
library(tidyverse)
```

## Issues

This document is to contain QA/QC issues that have come up during the merge efforts between CEDEN and SURF

### Load data

Start with all objects from **CEDENSURF_MERGE.rmd** in global environment (Open that rmd and choose "run all chunks")

Then remove all extra objects from working directory but CEDENSURF, which is the merged CEDEN and SURF dataset retaining data from CEDEN in SURF.

```{r}
# Load data

# original CEDEN WQ data

CEDEN_OriginalWQ <- fread("Data/CEDEN_WQ_20212259571.txt", quote = "") %>%
 filter(between(SampleDate, 
        as_date("2009-10-01"),as_date("2019-09-30"))) %>%
  mutate(Result = as.numeric(Result)) # Converts blanks to NA
 
# Ceden and SURF WQ data

SURFMod_WQ <- fread("https://github.com/WWU-IETC-R-Collab/CEDENSURF-mod/raw/main/Data/Output/SURFMod_water.csv")

CEDENMod_WQ <- fread("https://github.com/WWU-IETC-R-Collab/CEDEN-mod/raw/main/Data/Output/CEDENMod_WQ.csv")

# Merged output (as of 3/5/2021)

CEDENSURFNC <- fread("https://github.com/WWU-IETC-R-Collab/CEDENSURF-mod/raw/main/Data/Output/CEDENSURFMod.csv")
```

### 1. Non-results in CEDEN (NA) appear in SURF as "0" result concentration.

Conclusion: Our strategy to reduce CEDEN is removing valid records. 

This example was found when searching for causes of data sourced from CEDEN in SURF which was not retained in our CEDEN data - subsetting one station to identify the causes, and then going back to the originally downloaded CEDEN data, we were able to locate the records in question. 

**Glyphosate in SURF-CEDEN data, but not in CEDEN data.**

After our filtering and data mod, SURF-sourced data labeled as from CEDEN contains records of glyphosate measured at Grizzley Bay, yet the CEDEN set contains none including glyphosate.

An example record from SURF cited as coming from CEDEN:
```{r}
A <- CEDENSURF %>% filter(grepl('Grizzly', StationName)) %>%
  filter(grepl('Dolphin', StationName)) %>%
  filter(Source == "SURF") %>%
  filter(Data.source == "CEDEN")%>%
  filter(Analyte == "glyphosate")%>%
  select(Agency, Date, StationName, Analyte, Result)

A[1]
```

**Relocating the original data in CEDEN** confirmed that the record was not retained in our CEDEN result because it had NA results.

```{r}
C_OG <- CEDEN_OriginalWQ %>% filter(grepl('Grizzly', StationName)) %>%
  filter(grepl('Dolphin', StationName)) %>%
  filter(grepl('lyphos', Analyte))%>%
  select(SampleDate, StationName, Analyte, Result, SampleComments)

C_OG[1:4,]
C_OG$SampleComments[2]
```

**Interpretation?**

1. NA's could mean not tested, so it could be an error in SURF to retain those records, OR:

2. NA's in CEDEN's result column are sometimes paired with a "ND" code in the ResultQualCode - which suggests that they WERE tested for, and simply not detected. 

How many have a NA with no adjacent ResultQualCode can be found:

There are MANY! 262995 have NA results with ND codes. 509 have 0-results with ND codes, and the others have results that are numeric but extremely low.

```{r}
B <- CEDEN_OriginalWQ %>%
  filter(is.na(Result))

A<- CEDEN_OriginalWQ %>% filter(ResultQualCode == "ND")

Chk<- CEDEN_OriginalWQ %>% filter(ResultQualCode == "ND")  %>%
  filter(is.na(Result))

Chk<- CEDEN_OriginalWQ %>% filter(ResultQualCode == "ND")  %>%
  filter(Result == "0")
```

There are other analytes at this site recorded in SURF-CEDEN data but not in the data brought directly from CEDEN, too. While I have not investigated these yet, there is a chance they are indicative of this being a widespread issue.

```{r}
#Station Name: "Grizzly Bay at Dolphin nr. Suisun Slough. CEDEN: 207SNB0D7"

# Locate similar location records in CEDEN and SURF using flexible queries

A <- CEDENSURF %>% filter(grepl('Grizzly', StationName)) %>%
  filter(grepl('Dolphin', StationName)) %>%
  filter(Source == "SURF") %>%
  filter(Data.source == "CEDEN")

CS <- sort(unique(A$Analyte)) # 27

B <- CEDENSURF %>% filter(grepl('Grizzly', StationName)) %>%
  filter(grepl('Dolphin', StationName)) %>%
  filter(Source == "CEDEN")

C <- sort(unique(B$Analyte)) # 34

DIF<- setdiff(CS, C) #Analytes in ceden-surf and not in ceden

DIF
```

#### **2. CEDEN-SURF data not in CEDEN** 

Conclusion: Possibly indicative that data for certain areas of our study region may be omitted in the CEDEN data download. Alternatively, this project was incorrectly labeled as sourced from CEDEN.

SURF Name: "Toe Drain Nr Babel Slough Nr Freeport Ca"
SURF Code: "57_58"

Could not ID this location in CEDEN data, using grepl() to allow approximate naming. None with "Babel" and T in StationName

```{r}
# From SURF, identified as coming from CEDEN
CS <- CEDENSURF %>% filter(grepl('Toe', StationName)) %>%
  filter(grepl('Babel', StationName)) %>%
  filter(Data.source == "CEDEN")%>%
  filter(Source == "SURF")

# Direct from CEDEN
C <- CEDENSURF %>%
  filter(grepl('Babel', StationName)) %>%
  filter(Source == "CEDEN")
```

There are 657 records at this site from SURF, all from the same agency, and identified as coming from CEDEN. 

Here is one example record:

```{r}
nrow(CS)
unique(CS$Agency)
CS[1]
```
Also not in original water data...

```{r}
C_OG <- CEDEN_OriginalWQ %>%
  filter(grepl('Babel', StationName)) %>%
  select(SampleDate, StationName, Analyte, Result, ResultQualCode, SampleComments)
```

### 3. Naming inconsistencies between SURF and CEDEN

It appears that SURF has appended station names from CEDEN with extra info. Also, the methods to convert lat and long to geometry  For example:

SURF: "Grizzly Bay at Dolphin nr. Suisun Slough. CEDEN: 207SNB0D7"
SURF Code: "48_52"

CEDEN: "Grizzly Bay at Dolphin nr. Suisun Slough"
CEDEN Code: "207SNB0D7"

and 

SURF Name: "Sacramento River at Freeport (USGS-11447650)"
SURF Code: "34_5"

CEDEN Name: "Sacramento River at Freeport, CA"
CEDEN Code: 11447650

```{r}
A <- CEDENSURF %>% filter(grepl('Grizzly', StationName)) %>%
  filter(grepl('Dolphin', StationName)) %>%
  filter(Source == "SURF") %>%
  filter(Data.source == "CEDEN")%>%
  select(Agency, StationName, StationCode, Latitude, Longitude, geometry)

B <- CEDENSURF %>% filter(grepl('Grizzly', StationName)) %>%
  filter(grepl('Dolphin', StationName)) %>%
  filter(Source == "CEDEN")%>%
  select(Agency, StationName, StationCode, Latitude, Longitude, geometry)

rbind(A[1,], B[1,])
```

### **3. Other instances of data in SURF-CEDEN not in CEDEN, and vice versa**

If we subset just the data that occurs in CEDEN-SURF and not in our CEDEN set, we can hone in on some specific agencies and/or sites that account for the majority of the discrepancies.

```{r}
CS <- CEDENSURF %>%
  filter(Data.source == "CEDEN")%>%
  filter(Source == "SURF")

# Direct from CEDEN
C <- CEDENSURF %>%
  filter(Source == "CEDEN")

DIF<- setdiff(CS, C) #Analytes in ceden-surf and not in ceden

# Summarize by agency - sort descending (agency with the most in CS not in C at top)

SumDif <- DIF %>% group_by(Agency) %>%
  summarise(n = length(Latitude), 
            n_results = length(unique(Result))) %>%
  arrange(desc(n))

head(SumDif)
```
There are 93 stations represented.
```{r}
# Summarize by station - sort descending (agency with the most in CS not in C at top)

SumDif <- DIF %>% group_by(StationName) %>%
  summarise(n = length(Latitude), 
            n_results = length(unique(Result))) %>%
  arrange(desc(n))

head(SumDif)
```

**Let's take Ulatis Creek, n = 3,990**

Almost all the instances are with Result = 0; 3689/3990

```{r}
## Preview data, anything in common?

DIF %>%
  filter(grepl('Ulatis', StationName)) %>%
  filter(grepl('Creek', StationName)) %>%
  filter(grepl('Brown', StationName)) %>% 
  filter(Result == "0") %>% 
  nrow(.)
```

Subsetting from the merged dataset to see if the thousands of non-detects at this station are retained in CEDEN, we find that there are only 17 0-results at this station in the CEDEN data, compared with the CEDEN-SURF data.

```{r}
CS <- CEDENSURF %>%
  filter(Data.source == "CEDEN") %>%
  filter(Source == "SURF") %>%
  filter(grepl('Ulatis', StationName)) %>%
  filter(grepl('Creek', StationName)) %>%
  filter(grepl('Brown', StationName)) %>%
  filter(Result == "0")

C <- CEDENSURF %>%
  filter(Source == "CEDEN") %>%
  filter(grepl('Ulatis', StationName)) %>%
  filter(grepl('Creek', StationName))%>%
  filter(grepl('Brown', StationName))%>%
  filter(Result == "0")
```

In the original data at this site, there are 7492 NA results at this site, which were removed during our CEDEN data prep process. Within these, 7462 are ND qualcodes (indicating non-detect, or results below the MDL), and only 32 with NR qualcodes (which indicates the test failed)

```{r}
C_OG <- CEDEN_OriginalWQ %>% 
  filter(grepl('Ulatis', StationName)) %>%
  filter(grepl('Creek', StationName)) %>%
  filter(grepl('Brown', StationName)) %>%
  filter(is.na(Result)) %>%
  select(SampleDate, StationName, Analyte, Result, ResultQualCode, SampleComments)

head(C_OG)

# How many ND?

CEDEN_OriginalWQ %>% 
  filter(grepl('Ulatis', StationName)) %>%
  filter(grepl('Creek', StationName)) %>%
  filter(grepl('Brown', StationName)) %>%
  filter(is.na(Result)) %>%
  filter(ResultQualCode == "ND") %>%
  nrow(.)

# How many NR?

CEDEN_OriginalWQ %>% 
  filter(grepl('Ulatis', StationName)) %>%
  filter(grepl('Creek', StationName)) %>%
  filter(grepl('Brown', StationName)) %>%
  filter(is.na(Result)) %>%
  filter(ResultQualCode == "NR") %>%
  nrow(.)
```


#### 4. **Differences in geometry btwn SURF and CEDEN**

Same station at Grizzly Bay; despite same Lat and Long coordinates listed in SURF (1998 and 2012), compared to CEDEN (2010), the shapefile conversion to geometry is different.

```{r}
rbind(A[1,], B[1,])
```


#### 5. **Results in SURF vs CEDEN**

I'm interested, if SURF changed ND's in CEDEN from blank --> 0, would they also have changed values listed that were below MDL to 0? (Because theoretically also labeled ND)

Selected data in original set at one site, with qual code ND, and result above zero, then looked for that data in SURF by filtering on a specific analyte.

Prodiamine? No, also not found under value = 0 in C-S.
Penoxsulam? No, but YES in CS as result = 0. 
Fluridone?  Not in CEDENSURF with results 1-4, but YES in CEDENSURF 

*NOTE: Results in CEDEN are in ng/L whereas in CEDENSURF they're in PPB. Some results < MDL may be recorded as 0 in SURF*

**Example 1**
```{r}
# Subset records at one station with ND code and results above 0 (btwn 1-4)

C <- CEDEN_OriginalWQ %>% 
  filter(grepl('Ulatis', StationName)) %>%
  filter(grepl('Creek', StationName)) %>%
  filter(grepl('Brown', StationName)) %>%
  filter(ResultQualCode == "ND") %>%
  filter(between(Result,1,4) ) 

CSR <- CEDENSURF %>%
  filter(Data.source == "CEDEN") %>%
  filter(Source == "SURF") %>%
  filter(grepl('Ulatis', StationName)) %>%
  filter(grepl('Creek', StationName)) %>%
  filter(grepl('Brown', StationName)) %>%
  filter(between(Result,1,4) )

## Look for that ND record in SURF as ANY record with results in the appropriate range and with that analyte name. Not found. 

# Matches in Ceden OG
C2<- C %>%
  filter(grepl('Fluridone', Analyte)) %>%
  select(Agency = SubmittingAgency, StationName, Date = SampleDate, CollectionMethod = CollectionMethodName, Analyte, Result, Unit, ResultQualCode)

# Example: Delta RMP, on 2016-02-17
C2[2,]

# Matches in Ceden-sourced Surf data - none
CSR2<- CSR %>%
  filter(grepl('fluridone', Analyte))

# Matches in the CEDENSURF data allowing  another result value?
CS <- CEDENSURF %>%
  filter(Data.source == "CEDEN") %>%
  filter(Source == "SURF") %>%
  filter(grepl('Ulatis', StationName)) %>%
  filter(grepl('Creek', StationName)) %>%
  filter(grepl('Brown', StationName)) %>%
  filter(grepl('fluridone', Analyte))

CS2<- CS %>%
  filter(Date == "2016-02-17") %>%
  select(Agency, StationName, Date,CollectionMethod, Analyte, Result, Unit)

rbind(C2[2,], CS2, fill = TRUE)
```
This shows two samples that are likely duplicates, one "from ceden" in SURF, and the other from the original CEDEN file. The original CEDEN file has methods further described under "prep preservation name" = "lab filtered", which could explain the differences in collection method nomenclature.

**Example 2**

Lets try a new station with ND results - Freeport? It was actually really hard to find other examples! I may have been lucky with my first station/process.  Can come back to this. 

### Up next: Whether to retain or remove composite samples

If we decide to remove them, determine how to identify.

In CEDEN, might be able to locate by CollectionMethod OR CollectionDeviceDescription

ie: "7 day auto sampler" and "AutoSampler" collection methods may indicate composite over time, or "depth-integrating" collection device description may indivate composite over depths.
